### 2021-11-13 FinalProject445.Rmd
Added Project Rmd file

---
title: "DSCI-445 Project"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2021-11-27 gentestandtrn.R
Added heart.csv and generated test and training data sets with 60/40 split

@@ -1,0 +1,18 @@
+### Generating trainning and test data sets
+
+
+heart <- read.csv("heart.csv")
+
+library(tidyverse) 
+
+
+set.seed(445)
+n <- nrow(heart)
+trn <- seq_len(n) %in% sample(seq_len(n),round( 0.6*n))
+
+train <- heart[trn,]
+test <- heart[!trn,]
+
+
+write_csv(train, path = "train.csv")
+write_csv(test, path = "test.csv")

### 2021-11-27 eda-h.Rmd
created exploratory analysis file

---
title: "Exploratory Analysis"
author: "Henrique Magalhaes Rio"
output: html_notebook
---

```{r}

set.seed(445)
library(leaps)
library(ISLR)
library(tidyverse) 
library(knitr)
library(glmnet) # package for lasso and ridge regression
library(pls)
library(ISLR)
library(ggpubr)
library(MASS)
library(splines)
library(gam)
library(tree)
library(randomForest)
library(gbm)
library(caret)

```

```{r}
train <- read.csv("train.csv")
```

### 2021-11-27 eda-h.Rmd
started with ggpairs

@@
 library(caret)
+library(GGally)
 
@@
 train <- read.csv("train.csv")
+
+head(train)
+
+str(train)
 ```
+
+```{r}
+ggpairs(train)
+```

### 2021-11-27 gentestandtrn.R
fixed the structured of the data sets and re generated the csv files

@@ -1,7 +1,19 @@
-### Generating trainning and test data sets
+### Generating training and test data sets
 
 
 heart <- read.csv("heart.csv")
+
+
+heart$ChestPainType <- as.factor(heart$ChestPainType)
+heart$Sex <- as.factor(heart$Sex)
+heart$FastingBS <- as.factor(heart$FastingBS)
+heart$RestingECG <- as.factor(heart$RestingECG)
+heart$ExerciseAngina <- as.factor(heart$ExerciseAngina)
+heart$ST_Slope <- as.factor(heart$ST_Slope)
+heart$HeartDisease <- as.factor(heart$HeartDisease)
+
+str(heart)
+
 
 library(tidyverse) 
 
@@ -14,5 +26,7 @@ train <- heart[trn,]
 test <- heart[!trn,]
 
 
-write_csv(train, path = "train.csv")
-write_csv(test, path = "test.csv")
+write_csv(train, file  = "train.csv")
+write_csv(test, file = "test.csv")
+
+

### 2021-11-27 eda-h.Rmd
Creating some plots

@@
 
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+train$HeartDisease <- as.factor(train$HeartDisease)
+
 str(train)
@@
 ```{r}
-ggpairs(train)
+p1 <- ggplot(train,aes(x=Age,y=HeartDisease))+ geom_boxplot()+coord_flip()
+
+p2 <-ggplot(train,aes(x=Sex,fill=HeartDisease))+ geom_bar(position = "dodge")
+
+p3 <-ggplot(train,aes(x=ChestPainType,fill=HeartDisease))+ geom_bar(position = "dodge")
+
+p4<- ggplot(train,aes(x=RestingBP,y=HeartDisease))+ geom_boxplot()+coord_flip()
+
+ggarrange(p1,p2,p3,p4)
 ```
+
+```{r}
+p1 <- ggplot(train,aes(x=Cholesterol,y=HeartDisease))+ geom_boxplot()+coord_flip()
+
+p2 <-ggplot(train,aes(x=FastingBS,fill=HeartDisease))+ geom_bar(position = "dodge")
+
+p3 <-ggplot(train,aes(x=RestingECG,fill=HeartDisease))+ geom_bar(position = "dodge")
+
+p4<- ggplot(train,aes(x=MaxHR,y=HeartDisease))+ geom_boxplot()+coord_flip()
+
+ggarrange(p1,p2,p3,p4)
+```
+
+```{r}
+
+```

### 2021-11-27 FinalProject445.Rmd
Added Librarys
--

@@
 ```{r}
-plot(cars)
+set.seed(445)
+library(leaps)
+library(ISLR)
+library(tidyverse) 
+library(knitr)
+library(glmnet) # package for lasso and ridge regression
+library(pls)
+library(ISLR)
+library(ggpubr)
+library(MASS)
+library(splines)
+library(gam)
+library(tree)
+library(randomForest)
+library(gbm)
+library(caret)
 ```
-
-Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
-
-When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
-
-The preview shows you a rendered HTML copy of the contents of the editor.
-Consequently, unlike *Knit*, *Preview* does not run any R code chunks.
-Instead, the output of the chunk when it was last run in the editor is displayed.

### 2021-11-27 pca.R
first attempt on pca

@@ -1,0 +1,13 @@
+###
+
+
+
+train <- read.csv("train.csv")
+
+
+
+
+pca <- pca <- prcomp(train[,c("RestingBP","Cholesterol","HeartDisease","Oldpeak","MaxHR")], center = TRUE, scale = TRUE)
+
+
+biplot(pca)

### 2021-11-27 pca.R
fix

@@ -1,13 +1,22 @@
 ###
 
 
-
-train <- read.csv("train.csv")
-
+train <- read.csv("train.csv")head(train)
 
 
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+train$HeartDisease <- as.factor(train$HeartDisease)
 
-pca <- pca <- prcomp(train[,c("RestingBP","Cholesterol","HeartDisease","Oldpeak","MaxHR")], center = TRUE, scale = TRUE)
 
+
+pca <- prcomp(train[,c("RestingBP","Cholesterol","HeartDisease","Oldpeak","MaxHR")], center = TRUE, scale = TRUE)
+
+# ran into some problems, turn categorical variables into binary variable?
+pca2 <- prcomp(train,center = TRUE,scale = TRUE)
 
 biplot(pca)

### 2021-11-28 pca.R
Trying to get PCA working

@@ -1,22 +1,26 @@
 ###
 
+train <- read.csv("train.csv")
 
-train <- read.csv("train.csv")head(train)
 
 
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
-train$HeartDisease <- as.factor(train$HeartDisease)
+str(train)
+
+# try to make 
 
 
 
 pca <- prcomp(train[,c("RestingBP","Cholesterol","HeartDisease","Oldpeak","MaxHR")], center = TRUE, scale = TRUE)
 
-# ran into some problems, turn categorical variables into binary variable?
+# PCA does not work for categorical variabels.
 pca2 <- prcomp(train,center = TRUE,scale = TRUE)
 
-biplot(pca)
+
+# try to make dummy variables
+train$CPTASY <- ifelse(train$ChestPainType == 'ASY', 1, 0)
+
+str(train)
+
+pca2 <- prcomp(train[,c("RestingBP","Cholesterol","HeartDisease","Oldpeak","MaxHR","CPTASY")],center = TRUE,scale = TRUE)
+
+biplot(pca2)

### 2021-11-30 Bagging Model.Rmd
Adding Oriana's models (LASSO, PLS, Bagging)

---
title: "Bagging Model"
author: "Oriana Meldrum"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Bagging model

```{r}
#Set up
#reproducibility
set.seed(445)

library(ISLR)
library(tidyr)
library(dplyr)
library(knitr)
library(tree)
library(randomForest) # random forests & bagging

library(readr)
heart <- read_csv("heart.csv")
View(heart)

#60/40 split
heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

```

```{r}

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500) 
predict_bag <- predict(bagging, newdata = test)

mean((predict_bag - test$HeartDisease)^2)

#bagging
```

Test MSE is  0.1052255 for bagging

Now trying to visulize the bagging

```{r}

VI <- data.frame(var=names(heart[,-1]), imp=varImp(bagging))

#sort variable importance descending
VI_plot <- VI[order(VI$Overall, decreasing=TRUE),]

#visualize variable importance with horizontal bar plot
barplot(VI_plot$Overall,
        names.arg=rownames(VI_plot),
        horiz=TRUE,
        col='steelblue',
        xlab='Variable Importance')

```

### 2021-11-30 Lasso Model.Rmd
Adding Oriana's models (LASSO, PLS, Bagging)

---
title: "LASSO Model DSCI 445 Project"
author: "Oriana Meldrum"
date: "11/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lasso model: from hw 4

```{r}
#install.packages("glmnet")

#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls)

library(readr)
heart <- read_csv("heart.csv")
View(heart)

heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

```

```{r}
#60% training, 40% testing

train_matrix <- model.matrix(HeartDisease ~ ., data = train)
test_matrix <- model.matrix(HeartDisease ~ ., data = test)
grid <- 10^seq(10, -2, length=100)

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12)
plot(lasso)

#Here it breaks
#cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
#plot(cv_ridge)
#best_lambda_lasso <- cv_lasso$lambda.min

#lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
#lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

#num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

#values to use:
#best_lambda_lasso
#lasso_MSE
#num_zeros

```
Test MSE is  00000000000000000 for LASSO

```{r}
#lasso

x <- model.matrix(HeartDisease ~ . - 1, data = heart)
y <- heart$HeartDisease
lasso_heart <- cv.glmnet(x, y, type.measure = "mse")
plot(lasso_heart)

coef(lasso_heart)

sqrt(lasso_heart$cvm[lasso_heart$lambda == lasso_heart$lambda.1se])

```

```{r}

#cv_lasso <- cv.glmnet(heart, heart$HeartDisease, alpha = 1)
#plot(cv_lasso)
#plot(lasso_fit)

#opt_lambda <-cv_lasso$lambda.min
#opt_lambda

#fit.lasso <- glmnet(lasso_fit, Y, alpha = 1)
#predict(fit.lasso, s = opt_lambda, type = "coefficients")[1:11, ]

```

### 2021-11-30 PLS Model.Rmd
Adding Oriana's models (LASSO, PLS, Bagging)

---
title: "PLS Model"
author: "Oriana Meldrum"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#Set up
#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls)

library(readr)
heart <- read_csv("heart.csv")
View(heart)

#60/40 split
heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

```

Fitting a PLS model on the training set with $M$ chosen using 10-fold CV (on the training set only)

```{r}

pls_fit <- plsr(HeartDisease ~ ., data = train, scale = TRUE, validation = "CV")
validationplot(pls_fit, val.type = "MSEP")

pls_prediction <- predict(pls_fit, test, ncomp = 7)
pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)

m0 <- pcr(HeartDisease~.,data =train, scale = TRUE, validation = "CV")
mse <- MSEP(m0)

# 2. Create a plot of the CV MSE (note root MSE is reported) vs. $M$.
mse <- MSEP(m0)
data.frame(M = mse$comps, mse = t(as.data.frame(mse$val))[, "CV"]) %>%
  ggplot() +
  geom_line(aes(M, mse)) +
  geom_point(aes(M, mse))

m_pls <- mse$comps[which.min(as.data.frame(mse$val)[1,])]

#values to use:
mse
m_pls

```

Test MSE is  0.09841 for PLS at 13 variables

### 2021-11-30 eda-h.Rmd

@@
 
+When looking at age and HeartDisease there seems to be a clear difference between the two means , which can also be seem in the sex category.
+
 ```{r}
@@
 ```{r}
+p1 <- ggplot(train,aes(x=Oldpeak,y=HeartDisease))+ geom_boxplot()+coord_flip()
 
+p2 <-ggplot(train,aes(x=ExerciseAngina,fill=HeartDisease))+ geom_bar(position = "dodge")
+
+p3 <-ggplot(train,aes(x=ST_Slope,fill=HeartDisease))+ geom_bar(position = "dodge")
+
+ggarrange(p1,p2,p3)
 ```

### 2021-12-01 logit.Rmd
Logistic Regression File

---
title: "logit"
author: "Henrique Rio"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-12-01 logit.Rmd
Ran logistic model

@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(glmnet)
 ```
 
-## R Markdown
+```{r}
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+summary(model_glm)
 ```
 
-## Including Plots
+```{r}
+predicted2 <- predict(model_glm, newdata = test, "response")
 
-You can also embed plots, for example:
+predicted2 <- ifelse(predicted2>0.5,1,0)
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+confm2<-table(predicted=predicted2,True=test$HeartDisease)
+confm2
+paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
 ```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-12-05 presentation.Rmd
inital draft of presentation

---
title: "Group 9 Presentation"
author: "Henrique, Oriana, Jorie"
date: "12/4/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}

#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls) 
library(tree)
library(randomForest)
library(knitr)

train <- read.csv("train.csv")
test <- read.csv("test.csv")

```

## Group 9

Heart Disease
Oriana Meldrum, Henrique Rio, Jorie Alvis

## LDA

Jorie's LDA

## RF

Jorie's RF

## SVM

Jorie's SVM

## (the new thing)

Henrique's new thing insted of PCA

## Logistic Regression

Henrique's Logisic Regression

```{r}
model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
#summary(model_glm)

predicted2 <- predict(model_glm, newdata = test, "response")
predicted2 <- ifelse(predicted2>0.5,1,0)
confm2<-table(predicted=predicted2,True=test$HeartDisease)
kable(confm2)

paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
overall_log_reg <- (confm2[1,1]+confm2[2,2])/sum(confm2)
overall_log_reg

log_reg_mse <- mean((predicted2 - test$HeartDisease)^2) #MSE

```

## Boosting

Henrique's Boosting

## LASSO

Oriana's Lasso

```{r}

train_matrix <- model.matrix(HeartDisease ~ ., data = train)
test_matrix <- model.matrix(HeartDisease ~ ., data = test)
grid <- 10^seq(10, -2, length=100)

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
#plot(lasso)

cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min

lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

# MSE values to use:
best_lambda_lasso
lasso_MSE
num_zeros

#Confusion Matrix
#predicted_lasso <- predict(lasso, newdata = test, "response")
#predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
#confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
#confm_lasso

#Over all correct:
#overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)

#kable(confm_lasso) #confusion matrix
#overall_lasso #over all correct
```

## PLS

```{r}

pls_fit <- plsr(HeartDisease ~ ., data = train, type= "class", scale = TRUE, validation = "CV")
#validationplot(pls_fit, val.type = "MSEP")

pls_prediction <- predict(pls_fit, test,  ncomp = 7)
pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)

#Confusion Matrix prep
predicted_pls <- predict(pls_fit, newdata = test)
predicted_pls <- ifelse(predicted_pls>0.5,1,0)
#confm_pls<-table(predicted=predicted_pls,True=test$HeartDisease)

#Over all correct:

#overall_pls <- (confm_pls[1,1]+confm_pls[2,2])/sum(confm_pls)

m0 <- pcr(HeartDisease~.,data =train, scale = TRUE, validation = "CV")
mse <- MSEP(m0)

m_pls <- mse$comps[which.min(as.data.frame(mse$val)[1,])]

#values to use:
#mse
#m_pls

#kable(confm_pls) #confusion matrix
#overall_pls #over all correct

```

## Bagging
```{r}

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE) 
predict_bag <- predict(bagging, newdata = test)

#MSE
bagging_mse <- mean((predict_bag - test$HeartDisease)^2)

#Confusion Matrix
predicted_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predicted_bag>0.5,1,0)
confm_bag<-table(predicted=predicted_bag,True=test$HeartDisease)
confm_bag

#Over all correct:
overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)

kable(confm_bag) #confusion matrix
overall_bag #over all correct

```

## Comparision of all method's MSE

Chart with all MSE

```{r}
library(knitr)
df1 = data.frame(Method = c('LDA', 'RF', 'SVM', 'new THING', 'Logistic Regression', 'Boosting', 'LASSO', 'PLS', 'Bagging'),  MSE = c(0,0,0,0,0.1498638,0,0,0.1054 ,0.1165492))

kable(df1)

```

## Conclusion

the METHOD worked best, with an MSE of SOMETHING

```{r pressure}
plot(pressure)
```

### 2021-12-05 randomforest.Rmd
Random Forest Start

---
title: "RandomForest"
author: "Henrique Rio"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
train<-read.csv("train.csv")
test<-read.csv("test.csv")

```

```{r}
rf_fit <- randomForest(HeartDisease ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

testnmse <- mean((rf-true)^2)

paste("Random forest tree test error",testnmse)
```

### 2021-12-05 randomforest.Rmd
Got the forest working

@@
 ```{r}
-rf_fit <- randomForest(HeartDisease ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")
+rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")
 
@@
 
-testnmse <- mean((rf-true)^2)
+predicted2 <- ifelse(rf>0.5,1,0)
 
-paste("Random forest tree test error",testnmse)
+confm2<-table(predicted=rf,True=true)
+confm2
+
+paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
 ```
+
+```{r}
+print(rf_fit)
+```
+
+```{r}
+plot(rf_fit)
+```

### 2021-12-05 randomforest.Rmd
Representative Tree plot

@@
 
+options(repos='http://cran.rstudio.org')
+have.packages <- installed.packages()
+cran.packages <- c('devtools','plotrix','randomForest','tree')
+to.install <- setdiff(cran.packages, have.packages[,1])
+if(length(to.install)>0) install.packages(to.install)
+
+library(devtools)
+if(!('reprtree' %in% installed.packages())){
+  install_github('araastat/reprtree')
+}
+for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
+
+```
+
+```{r}
+library(reprtree)
 ```
@@
 ```
+
+```{r}
+reprtree:::plot.getTree(rf_fit)
+```
+
+https://stats.stackexchange.com/questions/41443/how-to-actually-plot-a-sample-tree-from-randomforestgettree
+
+Identifying representative trees from ensembles
+Banerjee, Mousumi ; Ding, Ying ; Noone, Anne-Michelle
+Chichester, UK: John Wiley & Sons, Ltd
+Statistics in medicine, 2012-07-10, Vol.31 (15), p.1601-1616

### 2021-12-05 randomforest.Rmd
factor modifications

@@
 library(reprtree)
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+train$HeartDisease <- as.factor(train$HeartDisease)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+test$HeartDisease <- as.factor(test$HeartDisease)
 ```

### 2021-12-05 logit.Rmd
some changes

@@
 
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+train$HeartDisease <- as.factor(train$HeartDisease)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+test$HeartDisease <- as.factor(test$HeartDisease)
+
 model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")

### 2021-12-05 presentation.Rmd
some changes

@@
 test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+train$HeartDisease <- as.factor(train$HeartDisease)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+test$HeartDisease <- as.factor(test$HeartDisease)
 
@@
 pls_prediction <- predict(pls_fit, test,  ncomp = 7)
-pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)
+#pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)
 

### 2021-12-05 presentation.Rmd
adding KNN model and fixing which model belongs to who

@@
 library(knitr)
+library(class)
+library(fastDummies)
 
@@
 
-Jorie's RF
+Henrique's RF
 
@@
 
-Henrique's Boosting
+Jories's Boosting
 
@@
 
-## PLS
+## KNN
 
@@
 
-pls_fit <- plsr(HeartDisease ~ ., data = train, type= "class", scale = TRUE, validation = "CV")
-#validationplot(pls_fit, val.type = "MSEP")
+temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
+temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))
 
-pls_prediction <- predict(pls_fit, test,  ncomp = 7)
-#pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)
+drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
+temp_train <- temp_train[, !(names(temp_train) %in% drops)]
+temp_test <-temp_test[, !(names(temp_test) %in% drops)]
 
-#Confusion Matrix prep
-predicted_pls <- predict(pls_fit, newdata = test)
-predicted_pls <- ifelse(predicted_pls>0.5,1,0)
-#confm_pls<-table(predicted=predicted_pls,True=test$HeartDisease)
+#trainknn <- temp_train[-temp_train$HeartDisease]
+#testknn <- test[-test$HeartDisease]
 
-#Over all correct:
+num_k <- seq(1:100)
 
-#overall_pls <- (confm_pls[1,1]+confm_pls[2,2])/sum(confm_pls)
+my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
 
-m0 <- pcr(HeartDisease~.,data =train, scale = TRUE, validation = "CV")
-mse <- MSEP(m0)
+knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)
 
-m_pls <- mse$comps[which.min(as.data.frame(mse$val)[1,])]
+confm_knn_1<-table(knn_pred_1,test$HeartDisease)
 
-#values to use:
-#mse
-#m_pls
+overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)
 
-#kable(confm_pls) #confusion matrix
-#overall_pls #over all correct
+overall_knn_1
+
+ #confusion matrix
+kable(confm_knn_1)
+mean(knn_pred_1 !=test$HeartDisease)
 

### 2021-12-06 presentation.Rmd
Added Random Forest and relative tree plot

@@
 knitr::opts_chunk$set(echo = FALSE)
+options(repos='http://cran.rstudio.org')
+have.packages <- installed.packages()
+cran.packages <- c('devtools','plotrix','randomForest','tree')
+to.install <- setdiff(cran.packages, have.packages[,1])
+if(length(to.install)>0) install.packages(to.install)
+
+library(devtools)
+if(!('reprtree' %in% installed.packages())){
+  install_github('araastat/reprtree')
+}
+for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
 ```
@@
 library(fastDummies)
+library(reprtree)
 
@@
 Henrique's RF
+
+```{r}
+rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")
+
+true <- test$HeartDisease
+rf <- predict(rf_fit,test,type="class")
+
+predicted2 <- ifelse(rf>0.5,1,0)
+
+confm2<-table(predicted=rf,True=true)
+kable(confm2)
+
+paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+
+rf_fit
+```
+
+## Representative Tree Paper
+
+```{r}
+reprtree:::plot.getTree(rf_fit)
+```
 

### 2021-12-06 LDA.Rmd
Added the LDA model
My LDA model had a test accuracy of 85.5% and MSE of 0.144. I am worried I did something wrong to get such a low MSE value.

---
title: "LDA_model"
author: "Jorie Alvis"
date: "December 1, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(MASS) # for lda

set.seed(445)
```

```{r Data, include=FALSE}
heart_train = read.csv("train.csv")

heart_train$Sex = factor(heart_train$Sex)
heart_train$ChestPainType = factor(heart_train$ChestPainType)
heart_train$FastingBS = factor(heart_train$FastingBS)
heart_train$RestingECG = factor(heart_train$RestingECG)
heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
heart_train$ST_Slope = factor(heart_train$ST_Slope)
heart_train$HeartDisease = factor(heart_train$HeartDisease)

heart_test = read.csv("test.csv")

heart_test$Sex = factor(heart_test$Sex)
heart_test$ChestPainType = factor(heart_test$ChestPainType)
heart_test$FastingBS = factor(heart_test$FastingBS)
heart_test$RestingECG = factor(heart_test$RestingECG)
heart_test$ExerciseAngina = factor(heart_test$ExerciseAngina)
heart_test$ST_Slope = factor(heart_test$ST_Slope)
heart_test$HeartDisease = factor(heart_test$HeartDisease)
```

```{r LDA}
LDA_model = lda(HeartDisease~., data=heart_train)

preds = predict(LDA_model, heart_train)
acc = sum(ifelse(preds$class == heart_train$HeartDisease, 1, 0))/nrow(heart_train) # 88.2%

table(preds$class, heart_train$HeartDisease)
acc
```

```{r Testing}
preds = predict(LDA_model, heart_test)
acc = sum(ifelse(preds$class == heart_test$HeartDisease, 1, 0))/nrow(heart_test) # 85.5%

MSE = sum((as.integer(heart_test$HeartDisease)-as.integer(preds$class))^2)/nrow(heart_test) # 0.1444

table(preds$class, heart_test$HeartDisease)
acc
MSE
```

The MSE seems to be way to low for our accuracy. (look at SVM accuracy and MSE)

### 2021-12-06 SVM.Rmd

---
title: "SVM_model"
author: "Jorie Alvis"
date: "December 1, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(e1071) # svm library

set.seed(445)
```

```{r Data, include=FALSE}
heart_train = read.csv("train.csv")

heart_train$Sex = factor(heart_train$Sex)
heart_train$ChestPainType = factor(heart_train$ChestPainType)
heart_train$FastingBS = factor(heart_train$FastingBS)
heart_train$RestingECG = factor(heart_train$RestingECG)
heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
heart_train$ST_Slope = factor(heart_train$ST_Slope)
heart_train$HeartDisease = factor(heart_train$HeartDisease)
```

```{r SVM_Linear}
# Train a Linear SVM and find the best cost value
# We set scale=TRUE becuase the continuous varibles seem to be normally distrabuted (from EDA)
svm_linear_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="linear", data=heart_train,
                      ranges = list(cost = seq(0.1, 5, by = 0.1)))
svm_linear = svm_linear_cv$best.model

preds = predict(svm_linear, heart_train, type="class")

table(heart_train$HeartDisease, preds)
```

```{r SVM_Polynomial}
# Train a Polynomial SVM
# Find the best cost and degree parameter
svm_poly_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="polynomial", data=heart_train,
                      ranges = list(cost = seq(0.1, 5, by = 0.1),
                                    degree = c(2,3,4, 5)))
svm_poly = svm_poly_cv$best.model

preds = predict(svm_poly, heart_train, type="class")

table(heart_train$HeartDisease, preds)
```

```{r SVM_Radial}
# Train a Radial SVM
# Find the best cost and gamma parameter
svm_radial_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="radial", data=heart_train,
                      ranges = list(cost = seq(0.2, 5, by = 0.2),
                                    gamma = seq(0.25, 3, by = 0.25)))
svm_radial = svm_radial_cv$best.model

preds = predict(svm_radial, heart_train, type="class")

table(heart_train$HeartDisease, preds)
```

```{r Testing}
heart_test = read.csv("test.csv")

preds_liner = predict(svm_linear, heart_test, type="class")
acc_liner = sum(ifelse(preds_liner == heart_test$HeartDisease, 1, 0))/nrow(heart_test)

preds_poly = predict(svm_poly, heart_test, type="class")
acc_poly = sum(ifelse(preds_poly == heart_test$HeartDisease, 1, 0))/nrow(heart_test)

preds_radial = predict(svm_radial, heart_test, type="class")
acc_radial = sum(ifelse(preds_radial == heart_test$HeartDisease, 1, 0))/nrow(heart_test) #87.2%

c("Linear Svm" = acc_liner, "Polynomial Svm" = acc_poly, "Radial SVM" = acc_radial)
```

The SVM kernal with the highest auccuracy was the radial kernal with a `r round(acc_radial, 3)*100`% accuracy. The Radial SVM had a cost parameter of `r svm_radial$cost` and gamma parameter of `r svm_radial$gamma`.

```{r Best_SVM}
table(heart_test$HeartDisease, preds_radial)
radial_mse = sum((heart_test$HeartDisease-as.integer(preds_radial))^2)/nrow(heart_test) #1.253
```

We can see from the confusion matrix of our Radial SVM that there is little bias in the miss-classifications while the MSE is `r round(radial_mse, 4)`.

### 2021-12-06 presentation.Rmd
representative tree fixe, new plot.
Fixed Presentation Representative tree

@@
 ```{r}
-rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")
+rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)-1), importance = TRUE, type= "classification")
 
@@
 
-## Representative Tree Paper
+## Representative Tree
 
 ```{r}
-reprtree:::plot.getTree(rf_fit)
+reptree <- ReprTree(rf_fit, train, metric='d2')
+plot(reptree, index=1)
+
 ```

### 2021-12-06 randomforest.Rmd
representative tree fixe, new plot.
Fixed Presentation Representative tree

@@
 
+set.seed(445)
+
+library(leaps)
+library(ISLR)
+library(tidyverse) 
+library(knitr)
+library(glmnet) # package for lasso and ridge regression
+library(pls)
+library(ISLR)
+library(ggpubr)
+library(MASS)
+library(splines)
+library(gam)
+library(tree)
+library(randomForest)
+library(gbm)
+library(caret)
 ```
@@
 ```{r}
-plot(rf_fit)
+reprtree:::plot.getTree(rf_fit)
 ```
@@
 ```{r}
-reprtree:::plot.getTree(rf_fit)
+
+reptree <- ReprTree(rf_fit, train, metric='d2')
+
 ```
+```{r}
+
+reptree$trees
+plot(reptree, index=1)
+
+```
+```{r}
+treet1 = predict(reptree, test, type = "class")
+
+table(predicted = treet1, actual = test$HeartDisease)
+```
+
+```{r}
+
+tree1 <- tree(HeartDisease~.,data= train)
+
+plot(tree1)
+text(tree1)
+
+tree
+
+```
+
+```{r}
+set.seed(445)
+
+treet = predict(tree1, test, type = "class")
+
+table(predicted = treet, actual = test$HeartDisease)
+
+```
+
+Overall test error rate (%83)
 

### 2021-12-06 presentation.Rmd
writing from representative trees

@@
 ## Representative Tree
+-Following the paper by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
+-This tree is calculated using a measure of similarity in which 2 trees are similar if they have the same prediction for all observations.
+-Using this idea on the random forest we get the following tree:
 
@@
 paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-overall_log_reg <- (confm2[1,1]+confm2[2,2])/sum(confm2)
-overall_log_reg
 
-log_reg_mse <- mean((predicted2 - test$HeartDisease)^2) #MSE
+ #MSE
 

### 2021-12-06 Bagging Model.Rmd
Updating models

@@
 library(readr)
-heart <- read_csv("heart.csv")
-View(heart)
 
 #60/40 split
-heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
-train <- heart[heart_df,]
-test <-  heart[-heart_df,]
+train <-train
+test <- test
 
@@
 ```{r}
+bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
+                        importance = TRUE,
+                        proximity = TRUE)
 
-bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500) 
+#Confusion Matrix
 predict_bag <- predict(bagging, newdata = test)
+predicted_bag <- ifelse(predict_bag>0.5,1,0)
+confm_bag<-table(predict_bag,test$HeartDisease)
+confm_bag
 
-mean((predict_bag - test$HeartDisease)^2)
+#MSE
+bagging_mse <- mean((predict_bag - test$HeartDisease)^2)
 
-#bagging
-```
+#Over all correct:
+overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
 
-Test MSE is  0.1052255 for bagging
+kable(confm_bag) #confusion matrix
+overall_bag #over all correct
 
-Now trying to visulize the bagging
-
-```{r}
-
-VI <- data.frame(var=names(heart[,-1]), imp=varImp(bagging))
-
+#Visulizaton of bagging
+VI <- data.frame(var=names(test[,-1]), imp=varImp(bagging))
+VI
 #sort variable importance descending
@@
 
+#bagging
 ```
+
+Test MSE is  0.1052255 for bagging

### 2021-12-06 KNN Model.Rmd
Updating models

--- 2021-12-06 presentation.Rmd
@@
 ---
-title: "Group 9 Presentation"
-author: "Henrique, Oriana, Jorie"
-date: "12/4/2021"
-output: ioslides_presentation
+title: "KNN Model"
+author: "Oriana Meldurm"
+date: "12/5/2021"
+output: pdf_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-options(repos='http://cran.rstudio.org')
-have.packages <- installed.packages()
-cran.packages <- c('devtools','plotrix','randomForest','tree')
-to.install <- setdiff(cran.packages, have.packages[,1])
-if(length(to.install)>0) install.packages(to.install)
-
-library(devtools)
-if(!('reprtree' %in% installed.packages())){
-  install_github('araastat/reprtree')
-}
-for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-```{r, include=FALSE}
-
-#reproducibility
-set.seed(445)
-
-library(ISLR)
-library(leaps)
-library(tidyr)
-library(ggplot2)
-library(glmnet)
-library(caret)
-library(pls) 
-library(tree)
-library(randomForest)
-library(knitr)
-library(class)
-library(fastDummies)
-library(reprtree)
-
-train <- read.csv("train.csv")
-test <- read.csv("test.csv")
-
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
-train$HeartDisease <- as.factor(train$HeartDisease)
-
-test$ChestPainType <- as.factor(test$ChestPainType)
-test$Sex <- as.factor(test$Sex)
-test$FastingBS <- as.factor(test$FastingBS)
-test$RestingECG <- as.factor(test$RestingECG)
-test$ExerciseAngina <- as.factor(test$ExerciseAngina)
-test$ST_Slope <- as.factor(test$ST_Slope)
-test$HeartDisease <- as.factor(test$HeartDisease)
-
-```
-
-## Group 9
-
-Heart Disease
-Oriana Meldrum, Henrique Rio, Jorie Alvis
-
-## LDA
-
-Jorie's LDA
-
-## RF
-
-Henrique's RF
-
-```{r}
-rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")
-
-true <- test$HeartDisease
-rf <- predict(rf_fit,test,type="class")
-
-predicted2 <- ifelse(rf>0.5,1,0)
-
-confm2<-table(predicted=rf,True=true)
-kable(confm2)
-
-paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-
-rf_fit
-```
-
-## Representative Tree Paper
-
-```{r}
-reprtree:::plot.getTree(rf_fit)
-```
-
-## SVM
-
-Jorie's SVM
-
-## (the new thing)
-
-Henrique's new thing insted of PCA
-
-## Logistic Regression
-
-Henrique's Logisic Regression
-
-```{r}
-model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
-#summary(model_glm)
-
-predicted2 <- predict(model_glm, newdata = test, "response")
-predicted2 <- ifelse(predicted2>0.5,1,0)
-confm2<-table(predicted=predicted2,True=test$HeartDisease)
-kable(confm2)
-
-paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-overall_log_reg <- (confm2[1,1]+confm2[2,2])/sum(confm2)
-overall_log_reg
-
-log_reg_mse <- mean((predicted2 - test$HeartDisease)^2) #MSE
-
-```
-
-## Boosting
-
-Jories's Boosting
-
-## LASSO
-
-Oriana's Lasso
+## R Markdown
 
@@
 
-train_matrix <- model.matrix(HeartDisease ~ ., data = train)
-test_matrix <- model.matrix(HeartDisease ~ ., data = test)
-grid <- 10^seq(10, -2, length=100)
+library(ISLR)
+library(class)
+library(tidyr)
 
-lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
-#plot(lasso)
-
-cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
-#plot(cv_lasso)
-best_lambda_lasso <- cv_lasso$lambda.min
-
-lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
-lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
-
-num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
-
-# MSE values to use:
-best_lambda_lasso
-lasso_MSE
-num_zeros
-
-#Confusion Matrix
-#predicted_lasso <- predict(lasso, newdata = test, "response")
-#predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
-#confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
-#confm_lasso
-
-#Over all correct:
-#overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
-
-#kable(confm_lasso) #confusion matrix
-#overall_lasso #over all correct
-```
-
-## KNN
-
-```{r}
+install.packages("fastDummies")
+library(fastDummies)
 
@@
 
-#trainknn <- temp_train[-temp_train$HeartDisease]
-#testknn <- test[-test$HeartDisease]
+trainknn <- temp_train[-temp_train$HeartDisease]
+testknn <- test[-test$HeartDisease]
 
+trainknn
+testknn
+
+?knn.cv
 num_k <- seq(1:100)
@@
 my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
+my_cv
 
@@
 
-overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)
+overall_knn_1 <- (confm_knn[1,1]+confm_knn[2,2])/sum(confm_knn)
 
@@
  #confusion matrix
-kable(confm_knn_1)
-mean(knn_pred_1 !=test$HeartDisease)
+kable(confm_knn)
+mean(knn_pred !=test$HeartDisease)
 
@@
 
-## Bagging
 ```{r}
 
-bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
-                        importance = TRUE,
-                        proximity = TRUE) 
-predict_bag <- predict(bagging, newdata = test)
+library(caret)
 
-#MSE
-bagging_mse <- mean((predict_bag - test$HeartDisease)^2)
+#5 fold
+trControl <- trainControl(method = "cv", number = 5)
 
-#Confusion Matrix
-predicted_bag <- predict(bagging, newdata = test)
-predicted_bag <- ifelse(predicted_bag>0.5,1,0)
-confm_bag<-table(predicted=predicted_bag,True=test$HeartDisease)
-confm_bag
+fit <- train(as.factor(HeartDisease) ~.,
+             method = "knn", 
+             tuneGrid = expand.grid(k = 1:100), 
+             trControl = trControl, 
+             metric = "Accuracy", 
+             data = temp_train)
 
-#Over all correct:
-overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
-
-kable(confm_bag) #confusion matrix
-overall_bag #over all correct
-
+fit
 ```
-
-## Comparision of all method's MSE
-
-Chart with all MSE
-
-```{r}
-library(knitr)
-df1 = data.frame(Method = c('LDA', 'RF', 'SVM', 'new THING', 'Logistic Regression', 'Boosting', 'LASSO', 'PLS', 'Bagging'),  MSE = c(0,0,0,0,0.1498638,0,0,0.1054 ,0.1165492))
-
-kable(df1)
-
-```
-
-## Conclusion
-
-the METHOD worked best, with an MSE of SOMETHING
-
-```{r pressure}
-plot(pressure)
-```

### 2021-12-06 Lasso Model.Rmd
Updating models

@@
 
-lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12)
-plot(lasso)
+lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
+#plot(lasso)
 
-#Here it breaks
-#cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
-#plot(cv_ridge)
-#best_lambda_lasso <- cv_lasso$lambda.min
+cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
+#plot(cv_lasso)
+best_lambda_lasso <- cv_lasso$lambda.min
 
-#lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
-#lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
+lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
+lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
 
-#num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
+num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
 
-#values to use:
-#best_lambda_lasso
-#lasso_MSE
-#num_zeros
+# MSE values to use:
+best_lambda_lasso
+lasso_MSE
+num_zeros
+
+#Confusion Matrix
+predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
+predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
+confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
+confm_lasso
+
+#Over all correct:
+overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
+
+kable(confm_lasso) #confusion matrix
+overall_lasso #over all correct
 

### 2021-12-06 PLS Model.Rmd
Updating models

@@
 Test MSE is  0.09841 for PLS at 13 variables
+
+## PLS
+
+```{r}
+
+pls_fit <- plsr(HeartDisease ~ ., data = train, type= "class", scale = TRUE, validation = "CV")
+#validationplot(pls_fit, val.type = "MSEP")
+
+pls_prediction <- predict(pls_fit, test,  ncomp = 7)
+#pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)
+
+#Confusion Matrix prep
+predicted_pls <- predict(pls_fit, newdata = test)
+predicted_pls <- ifelse(predicted_pls>0.5,1,0)
+#confm_pls<-table(predicted=predicted_pls,True=test$HeartDisease)
+
+#Over all correct:
+
+#overall_pls <- (confm_pls[1,1]+confm_pls[2,2])/sum(confm_pls)
+
+m0 <- pcr(HeartDisease~.,data =train, scale = TRUE, validation = "CV")
+mse <- MSEP(m0)
+
+m_pls <- mse$comps[which.min(as.data.frame(mse$val)[1,])]
+
+#values to use:
+#mse
+#m_pls
+
+#kable(confm_pls) #confusion matrix
+#overall_pls #over all correct
+
+```

### 2021-12-06 presentation.Rmd
Commit
Updating models

@@
 date: "12/4/2021"
-output: ioslides_presentation
+output:
+  beamer_presentation: default
+  ioslides_presentation: default
 ---
@@
 library(tidyr)
+library(dplyr)
 library(ggplot2)
@@
 Oriana Meldrum, Henrique Rio, Jorie Alvis
+
+## Introduction
+
+In 2019 659,041 Americans died of heart disease.
+Four out of five heart disease deaths are due to heart attacks and strokes, and one-third of these deaths occur prematurely in people under 70 years of age.
+Heart failure is a common event caused by heart disease and this dataset we used contains 11 variables that can be used to predict a possible heart disease.
+
+People with cardiovascular disease or who are at high cardiovascular risk (due to the presence of one or more risk factors) need early detection and management so machine learning techniques are of great clinical intrest for early detection.
+
+(https://www.cdc.gov/nchs/fastats/leading-causes-of-death.htm, https://www.kaggle.com/fedesoriano/heart-failure-prediction)
 
@@
 
-paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
 
@@
 
-cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
+cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
 #plot(cv_lasso)
@@
 #Confusion Matrix
-#predicted_lasso <- predict(lasso, newdata = test, "response")
-#predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
-#confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
-#confm_lasso
+predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
+predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
+confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
+confm_lasso
 
 #Over all correct:
-#overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
+overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
 
-#kable(confm_lasso) #confusion matrix
-#overall_lasso #over all correct
+kable(confm_lasso) #confusion matrix
+overall_lasso #over all correct
 ```
@@
 mean(knn_pred_1 !=test$HeartDisease)
-
 ```
@@
                         proximity = TRUE)
+
+#Confusion Matrix
 predict_bag <- predict(bagging, newdata = test)
+predicted_bag <- ifelse(predict_bag>0.5,1,0)
+confm_bag<-table(predict_bag,test$HeartDisease)
+confm_bag
 
@@
 bagging_mse <- mean((predict_bag - test$HeartDisease)^2)
-
-#Confusion Matrix
-predicted_bag <- predict(bagging, newdata = test)
-predicted_bag <- ifelse(predicted_bag>0.5,1,0)
-confm_bag<-table(predicted=predicted_bag,True=test$HeartDisease)
-confm_bag
 
@@
 overall_bag #over all correct
+
+#Visulizaton of bagging
+VI <- data.frame(var=names(test[,-1]), imp=varImp(bagging))
+
+#sort variable importance descending
+VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
+
+#visualize variable importance with horizontal bar plot
+barplot(VI_plot$imp.0,
+        names.arg=rownames(VI_plot),
+        horiz=TRUE,
+        col='steelblue',
+        xlab='Variable Importance')
 
@@
 the METHOD worked best, with an MSE of SOMETHING
-
-```{r pressure}
-plot(pressure)
-```

### 2021-12-07 presentation.Rmd
New Slides

@@
 output:
+  slidy_presentation: default
+  ioslides_presentation: default
   beamer_presentation: default
-  ioslides_presentation: default
 ---

### 2021-12-07 presentation.Rmd
Done with rf v1

@@
 
+## Data
+
+- There are 918 observations which we split into 60/40 for training and validation.
+
+- 12 predictors are included in the data which are:
+
++ Age
+
++ Sex M/F
+
++ Chest Pain Type:
+
+++ TA: Typical Angina
+
+++ ATA: Atypical Angina
+
+++ NAP: Non-Anginal Pain
+
+++ ASY: Asymptomatic
+
++ Resting Blood Pressure [mm Hg]
+
++ Cholesterol [mm/dl]
+
++ Fasting Blood Sugar Y/N
+
++ Resting Electrocardiogram
+
++ Normal: Normal
+
+++ ST: Having ST-T Wave abnormality
+
+++ LVH: Showing probable or definite left ventricular hypertrophy.
+
++ Max HR: Maximum heart rate achieved.
+
++ Exercise Angina: exercise-induced angine Y/N
+
++ Old peak: Oldpeak= ST value of depression
+
++ ST_Slope:
+
+++ Up sloping(up)
+
+++ Flat(flat)
+
+++ Down sloping (down)
+
++ Heart Disease: [1:Heart Disease, 0: Normal]
+
 ## LDA
@@
 rf_fit
+
+imp<-data.frame(rf_fit$importance)
+
+imp$variables <- rownames(imp)
+rownames(imp) <- 1:nrow(imp)
+
+imp <- imp[order(imp$MeanDecreaseAccuracy),]
+
+ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())
 ```
@@
 ## Representative Tree
--Following the paper by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
--This tree is calculated using a measure of similarity in which 2 trees are similar if they have the same prediction for all observations.
--Using this idea on the random forest we get the following tree:
+- Following the paper by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
+- This tree is calculated using a measure of similarity in which 2 trees are similar if they have the same prediction for all observations.
+- Using this idea on the random forest we get the following tree:
 
-```{r}
+```{r, fig.height=5,fig.width=10}
 reptree <- ReprTree(rf_fit, train, metric='d2')

### 2021-12-07 boosting.Rmd
boosting file

---
title: "Boosting"
author: "Henrique"
output: html_notebook
---

```{r}

train <- read.csv("train.csv")
test <- read.csv("test.csv")
library(gbm)
library(caret)

```

```{r}

train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)

test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)

set.seed(445)
lambda=seq(from=0,to=1,length.out=100)

tmse<-0

boost <- gbm(HeartDisease~.,data=temp_train,n.trees=1000,shrinkage=4,interaction.depth = 2 ,distribution = "bernoulli")

gbm_grid =  expand.grid(interaction.depth = 1:5,
                        n.trees = (1:6) * 500,
                        shrinkage = c(0.001, 0.01, 0.1),
                        n.minobsinnode = 10)

```

### 2021-12-07 EDA-J.Rmd
Created EDA markdown
Created an R markdown for EDA for our training dataset

---
title: "EDA-J"
author: "Jorie Alvis"
date: "December 7, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

set.seed(445)

heart_train = read.csv("train.csv")
```

```{r Cor}
pairs(heart_train)
pairs(heart_train[,-c(2,3,6,7,9,10,11)])
```

```{r Norm, echo=FALSE}
qqnorm(heart_train$Age, main="QQ Plot of Age")
qqline(heart_train$Age)

qqnorm(heart_train$RestingBP, main="QQ Plot of Resting BP")
qqline(heart_train$RestingBP)

qqnorm(heart_train$Cholesterol, main="QQ Plot of Cholesterol")
qqline(heart_train$Cholesterol)

qqnorm(heart_train$MaxHR, main="QQ Plot of Max HR")
qqline(heart_train$MaxHR)
```

### 2021-12-07 Boosting-J.Rmd
Renamed Jorie's boosting model file
Renamed my boosting model R file to not conflict with Henrique's

---
title: "Boosting"
author: "Jorie Alvis"
date: "December 5, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(gbm) # boosting package

set.seed(445)
```

```{r Data, include=FALSE}
heart_train = read.csv("train.csv")

heart_train$Sex = factor(heart_train$Sex)
heart_train$ChestPainType = factor(heart_train$ChestPainType)
heart_train$FastingBS = factor(heart_train$FastingBS)
heart_train$RestingECG = factor(heart_train$RestingECG)
heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
heart_train$ST_Slope = factor(heart_train$ST_Slope)
#heart_train$HeartDisease = factor(heart_train$HeartDisease)
heart_train$HeartDisease = as.integer(heart_train$HeartDisease)

heart_test = read.csv("test.csv")

heart_test$Sex = factor(heart_test$Sex)
heart_test$ChestPainType = factor(heart_test$ChestPainType)
heart_test$FastingBS = factor(heart_test$FastingBS)
heart_test$RestingECG = factor(heart_test$RestingECG)
heart_test$ExerciseAngina = factor(heart_test$ExerciseAngina)
heart_test$ST_Slope = factor(heart_test$ST_Slope)
#heart_test$HeartDisease = factor(heart_test$HeartDisease)
heart_test$HeartDisease = as.integer(heart_test$HeartDisease)
```

```{r Boosting}
# The response needs to be an integer not a factor
boosted_model = gbm(HeartDisease~., data=heart_train, n.trees=10000, interaction.depth=3,
                    shrinkage=0.01, distribution="bernoulli", cv.folds=10)
gbm.perf(boosted_model, plot.it = TRUE, method="cv")
```

```{r Testing}
t=gbm(HeartDisease~., data=heart_train, n.trees=1000, interaction.depth=3,
                    shrinkage=0.075, distribution="bernoulli")

preds = predict(t, heart_test, type="response")
preds2 = ifelse(preds<0.5, 0, 1)

acc = sum(as.integer(preds2==heart_test$HeartDisease))/nrow(heart_test)

table(preds2, "True"=heart_test$HeartDisease)
acc
summary(t)
```

### 2021-12-07 presentation.Rmd
CHANGES

@@
 
-## Data
+## Data {.smaller}
 
@@
 
+## Data {.smaller}
+
 + Resting Electrocardiogram
 
-+ Normal: Normal
+++ Normal: Normal
 
@@
 
-## RF
-
-Henrique's RF
+## Random Forest
 
@@
 
-ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())
+```
+
+## Random Forest
+
+- Variable importance shows.
+
+```{r}
+ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())
 ```
@@
 
-```{r, fig.height=5,fig.width=10}
+## Tree
+```{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
 reptree <- ReprTree(rf_fit, train, metric='d2')
+
 plot(reptree, index=1)
@@
 ## Bagging
-```{r}
+```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
 
@@
 #Visulizaton of bagging
-VI <- data.frame(var=names(test[,-1]), imp=varImp(bagging))
+VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
 
@@
 #visualize variable importance with horizontal bar plot
-barplot(VI_plot$imp.0,
-        names.arg=rownames(VI_plot),
-        horiz=TRUE,
-        col='steelblue',
-        xlab='Variable Importance')
+#barplot(VI_plot$imp.0,
+   #     names.arg=rownames(VI_plot),
+    #    horiz=TRUE,
+     #   col='steelblue',
+      #  xlab='Variable Importance')
+
+ggplot(VI_plot)+geom_col(aes(x=(rownames(VI_plot)),y=imp.0,fill=rownames(VI_plot)))+coord_flip()+guides(fill=FALSE)+
+  labs(x="Variable Name",y="Importance")
 

### 2021-12-07 boosting.Rmd
Fixed file name error
Henrique's Boosting

@@
 
-boost <- gbm(HeartDisease~.,data=temp_train,n.trees=1000,shrinkage=4,interaction.depth = 2 ,distribution = "bernoulli")
+oob = trainControl(method = "oob")
+cv_5 = trainControl(method = "cv", number = 5)
 
@@
 
+seat_gbm_tune = train(as.factor(HeartDisease) ~ ., data = train,
+                      method = "gbm",
+                      trControl = cv_5,
+                      verbose = FALSE,
+                      tuneGrid = gbm_grid)
+
+plot(seat_gbm_tune)
+
 ```
+
+```{r}
+boost <- gbm(HeartDisease~.,data=train,n.trees=1000,shrinkage=0.01,interaction.depth = 1 ,distribution = "bernoulli")
+
+boostpred = ifelse(predict(boost,test, n.trees = 2000, "response") > 0.5, 
+                             "1", "0")
+
+confm2<-table(predicted = boostpred, actual = test$HeartDisease)
+confm2
+error<-(confm2[1,1]+confm2[2,2])/sum(confm2)
+error
+```

### 2021-12-07 boosting.Rmd
Added Boosting importance

@@
 library(caret)
+
+set.seed(445)
 
@@
 ```{r}
-boost <- gbm(HeartDisease~.,data=train,n.trees=1000,shrinkage=0.01,interaction.depth = 1 ,distribution = "bernoulli")
+boost <- gbm(HeartDisease~.,data=train,n.trees=500,shrinkage=0.01,interaction.depth = 1 ,distribution = "bernoulli")
 
-boostpred = ifelse(predict(boost,test, n.trees = 2000, "response") > 0.5, 
+boostpred = ifelse(predict(boost,test, n.trees = 500, "response") > 0.5, 
                              "1", "0")
@@
 ```
+
+```{r}
+
+summary(boost)
+
+```

### 2021-12-07 presentation.Rmd
Added my LDA and SVM code
Added the code for my LDA and SVM models into the presentation

@@
 
+```{r LDA}
+LDA_model = lda(HeartDisease~., data=train)
+
+preds = predict(LDA_model, test)
+
+acc = sum(ifelse(preds$class == test$HeartDisease, 1, 0))/nrow(test)
+
+table("Preds"=preds$class, "True"=test$HeartDisease)
+```
+
 ## Random Forest
@@
 Jorie's SVM
+
+```{r SVM}
+SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
+
+preds = predict(SVM_radial, test, type="class")
+acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
+
+table(test$HeartDisease, preds)
+```
 

### 2021-12-07 presentation.Rmd
Presentation File/ Changes to LDA

@@
 library(reprtree)
+library(MASS)
+library(e1071)
 
@@
 table("Preds"=preds$class, "True"=test$HeartDisease)
+
+overall_LDA <- acc
+overall_LDA
+
+LDA_model
 ```

### 2021-12-08 boosting.Rmd
Lot of formatting on the presentation, removed last slide into one Results slide, added boostings, bagging is now broken.

@@
+@@ -1,70 +0,0 @@
 ---
@@
 set.seed(445)
-lambda=seq(from=0,to=1,length.out=100)
-
-tmse<-0
 
@@
 gbm_grid =  expand.grid(interaction.depth = 1:5,
-                        n.trees = (1:6) * 500,
-                        shrinkage = c(0.001, 0.01, 0.1),
+                        n.trees = (1:5) * 200,
+                        shrinkage = c(0.01,0.015,0.1),
                         n.minobsinnode = 10)
@@
 ```{r}
-boost <- gbm(HeartDisease~.,data=train,n.trees=500,shrinkage=0.01,interaction.depth = 1 ,distribution = "bernoulli")
+boost <- gbm(HeartDisease~.,data=train,n.trees=600,shrinkage=0.01,interaction.depth = 2 ,distribution = "bernoulli")
 
@@
 ```{r}
+summa<-summary(boost)
+```
 
-summary(boost)
-
+```{r}
+ggplot(summa)+geom_col(aes(x=var,y=rel.inf,fill=var))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
 ```

### 2021-12-08 presentation.Rmd
Lot of formatting on the presentation, removed last slide into one Results slide, added boostings, bagging is now broken.

@@
 library(e1071)
+library(gbm)
 
@@
 
-Jorie's LDA
-
 ```{r LDA}
@@
 
-predicted2 <- ifelse(rf>0.5,1,0)
-
 confm2<-table(predicted=rf,True=true)
@@
 
+overall_RF <- (confm2[1,1]+confm2[2,2])/sum(confm2)
 rf_fit
@@
 
-- Variable importance shows.
+- ST_Slope
+-
 
 ```{r}
-ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())
+ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
 ```
@@
 ## Representative Tree
-- Following the paper by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
-- This tree is calculated using a measure of similarity in which 2 trees are similar if they have the same prediction for all observations.
+- Following the paper "Identifying representative Trees From Ensembles"  by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
+
+- This tree is calculated using a measure of similarity, in which 2 trees are similar if they have the same prediction for all observations.
+
+- The representative tree is then choosen based on its averaged similarity between all other trees.
+
 - Using this idea on the random forest we get the following tree:
 
-## Tree
+## Representative Tree Plot
 ```{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
@@
 
-Jorie's SVM
-
 ```{r SVM}
@@
 
-table(test$HeartDisease, preds)
+table(predicted=preds,true=test$HeartDisease)
+
+overall_SVM <- acc
+
+paste("Accuracy", acc)
 ```
-
-## (the new thing)
-
-Henrique's new thing insted of PCA
 
@@
 
-Henrique's Logisic Regression
-
 ```{r}
 model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
-#summary(model_glm)
+summary(model_glm)
 
@@
 
- #MSE
+overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
 
@@
 
-Jories's Boosting
+```{r}
+
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+
+set.seed(445)
+
+oob = trainControl(method = "oob")
+cv_5 = trainControl(method = "cv", number = 5)
+
+gbm_grid =  expand.grid(interaction.depth = 1:5,
+                        n.trees = (1:5) * 200,
+                        shrinkage = c(0.01,0.015,0.1),
+                        n.minobsinnode = 10)
+
+gbm_tune = train(as.factor(HeartDisease) ~ ., data = train,
+                      method = "gbm",
+                      trControl = cv_5,
+                      verbose = FALSE,
+                      tuneGrid = gbm_grid)
+
+plot(gbm_tune)
+
+```
+
+```{r}
+boost <- gbm(HeartDisease~.,data=train,n.trees=600,shrinkage=0.01,interaction.depth = 2 ,distribution = "bernoulli")
+
+boostpred = ifelse(predict(boost,test, n.trees = 500, "response") > 0.5, "1", "0")
+
+confboost<-table(predicted = boostpred, actual = test$HeartDisease)
+confboost
+error<-(confboost[1,1]+confboost[2,2])/sum(confboost)
+overall_boost <- error
+```
+
+```{r,include=FALSE}
+summa<-summary(boost)
+```
+
+```{r}
+ggplot(summa)+geom_col(aes(x=var,y=rel.inf,fill=var))+coord_flip()+labs(y="Relative Influence", x= " Variable")+guides(fill="none")
+```
 
@@
 ```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
 
@@
 confm_bag<-table(predict_bag,test$HeartDisease)
-confm_bag
 
@@
 #sort variable importance descending
-VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
+#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
 
@@
 
-ggplot(VI_plot)+geom_col(aes(x=(rownames(VI_plot)),y=imp.0,fill=rownames(VI_plot)))+coord_flip()+guides(fill=FALSE)+
+ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill=FALSE)+
   labs(x="Variable Name",y="Importance")
-
 ```
 
-## Comparision of all method's MSE
+## Results
 
-Chart with all MSE
+- What method should be choosen?
 
@@
 library(knitr)
-df1 = data.frame(Method = c('LDA', 'RF', 'SVM', 'new THING', 'Logistic Regression', 'Boosting', 'LASSO', 'PLS', 'Bagging'),  MSE = c(0,0,0,0,0.1498638,0,0,0.1054 ,0.1165492))
+df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  MSE = c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
 
@@
 ```
-
-## Conclusion
-
-the METHOD worked best, with an MSE of SOMETHING

### 2021-12-08 presentation.Rmd
BAGGING FIXED NOW, SVM IS BROKEN

@@
 title: "Group 9 Presentation"
-author: "Henrique, Oriana, Jorie"
-date: "12/4/2021"
+subtitle: "Heart Disease Classification"
+author: "Henrique Magalhaes Rio , Oriana Meldrum, Jorie Alvis"
+date: "12/8/2021"
 output:
@@
 
-## Group 9
-
-Heart Disease
-Oriana Meldrum, Henrique Rio, Jorie Alvis
-
 ## Introduction
@@
 
+## Logistic Regression
+
+```{r}
+model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
+summary(model_glm)
+
+predicted2 <- predict(model_glm, newdata = test, "response")
+predicted2 <- ifelse(predicted2>0.5,1,0)
+confm2<-table(predicted=predicted2,True=test$HeartDisease)
+kable(confm2)
+
+paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+
+overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
+
+```
+
 ## LDA
@@
 ```
+## KNN
+
+```{r}
+
+temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
+temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))
+
+drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
+temp_train <- temp_train[, !(names(temp_train) %in% drops)]
+temp_test <-temp_test[, !(names(temp_test) %in% drops)]
+
+#trainknn <- temp_train[-temp_train$HeartDisease]
+#testknn <- test[-test$HeartDisease]
+
+num_k <- seq(1:100)
+
+my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
+
+knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)
+
+confm_knn_1<-table(knn_pred_1,test$HeartDisease)
+
+overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)
+
+overall_knn_1
+
+ #confusion matrix
+kable(confm_knn_1)
+mean(knn_pred_1 !=test$HeartDisease)
+```
+
+## LASSO
+
+```{r}
+
+train_matrix <- model.matrix(HeartDisease ~ ., data = train)
+test_matrix <- model.matrix(HeartDisease ~ ., data = test)
+grid <- 10^seq(10, -2, length=100)
+
+lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
+#plot(lasso)
+
+cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
+#plot(cv_lasso)
+best_lambda_lasso <- cv_lasso$lambda.min
+
+lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
+lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
+
+num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
+
+# MSE values to use:
+best_lambda_lasso
+lasso_MSE
+num_zeros
+
+#Confusion Matrix
+predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
+predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
+confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
+confm_lasso
+
+#Over all correct:
+overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
+
+kable(confm_lasso) #confusion matrix
+overall_lasso #over all correct
+```
+
+## Bagging
+```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+
+bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
+                        importance = TRUE,
+                        proximity = TRUE)
+
+#Confusion Matrix
+predict_bag <- predict(bagging, newdata = test)
+predicted_bag <- ifelse(predict_bag>0.5,1,0)
+confm_bag<-table(predict_bag,test$HeartDisease,)
+
+#MSE
+bagging_mse <- mean((predict_bag - test$HeartDisease)^2)
+
+#Over all correct:
+overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
+
+kable(confm_bag) #confusion matrix
+overall_bag #over all correct
+
+#Visulizaton of bagging
+VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
+
+#sort variable importance descending
+#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
+
+#visualize variable importance with horizontal bar plot
+#barplot(VI_plot$imp.0,
+   #     names.arg=rownames(VI_plot),
+    #    horiz=TRUE,
+     #   col='steelblue',
+      #  xlab='Variable Importance')
+
+ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill=FALSE)+
+  labs(x="Variable Name",y="Importance")
+```
 
@@
 confm2<-table(predicted=rf,True=true)
-kable(confm2)
+confm2
 
@@
 
-## Random Forest
-
-- ST_Slope
--
+## Random Forest Plot
 
@@
 plot(reptree, index=1)
-
-```
-
-## SVM
-
-```{r SVM}
-SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
-
-preds = predict(SVM_radial, test, type="class")
-acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
-
-table(predicted=preds,true=test$HeartDisease)
-
-overall_SVM <- acc
-
-paste("Accuracy", acc)
-```
-
-## Logistic Regression
-
-```{r}
-model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
-summary(model_glm)
-
-predicted2 <- predict(model_glm, newdata = test, "response")
-predicted2 <- ifelse(predicted2>0.5,1,0)
-confm2<-table(predicted=predicted2,True=test$HeartDisease)
-kable(confm2)
-
-paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-
-overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
 
@@
 
+- Tuning Parmeters Choosen= 600 trees, shrinkage=0.01, interaction= 2.
+
 ```{r}
@@
 
-## LASSO
+## SVM
 
-Oriana's Lasso
+```{r SVM}
 
-```{r}
-
-train_matrix <- model.matrix(HeartDisease ~ ., data = train)
-test_matrix <- model.matrix(HeartDisease ~ ., data = test)
-grid <- 10^seq(10, -2, length=100)
-
-lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
-#plot(lasso)
-
-cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
-#plot(cv_lasso)
-best_lambda_lasso <- cv_lasso$lambda.min
-
-lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
-lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
-
-num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
-
-# MSE values to use:
-best_lambda_lasso
-lasso_MSE
-num_zeros
-
-#Confusion Matrix
-predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
-predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
-confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
-confm_lasso
-
-#Over all correct:
-overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
-
-kable(confm_lasso) #confusion matrix
-overall_lasso #over all correct
-```
-
-## KNN
-
-```{r}
-
-temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
-temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))
-
-drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
-temp_train <- temp_train[, !(names(temp_train) %in% drops)]
-temp_test <-temp_test[, !(names(temp_test) %in% drops)]
-
-#trainknn <- temp_train[-temp_train$HeartDisease]
-#testknn <- test[-test$HeartDisease]
-
-num_k <- seq(1:100)
-
-my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
-
-knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)
-
-confm_knn_1<-table(knn_pred_1,test$HeartDisease)
-
-overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)
-
-overall_knn_1
-
- #confusion matrix
-kable(confm_knn_1)
-mean(knn_pred_1 !=test$HeartDisease)
-```
-
-## Bagging
-```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
 train <- read.csv("train.csv")
@@
 
-bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
-                        importance = TRUE,
-                        proximity = TRUE)
+SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
 
-#Confusion Matrix
-predict_bag <- predict(bagging, newdata = test)
-predicted_bag <- ifelse(predict_bag>0.5,1,0)
-confm_bag<-table(predict_bag,test$HeartDisease)
+preds = predict(SVM_radial, test, type="class")
+acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
 
-#MSE
-bagging_mse <- mean((predict_bag - test$HeartDisease)^2)
+table(preds,test$HeartDisease)
 
-#Over all correct:
-overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
+overall_SVM <- acc
 
-kable(confm_bag) #confusion matrix
-overall_bag #over all correct
-
-#Visulizaton of bagging
-VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
-
-#sort variable importance descending
-#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
-
-#visualize variable importance with horizontal bar plot
-#barplot(VI_plot$imp.0,
-   #     names.arg=rownames(VI_plot),
-    #    horiz=TRUE,
-     #   col='steelblue',
-      #  xlab='Variable Importance')
-
-ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill=FALSE)+
-  labs(x="Variable Name",y="Importance")
+paste("the overall fraction of correct predictions in the test dataset is :", acc)
 ```
@@
 
-- What method should be choosen?
+- What method would be the best?
 

### 2021-12-08 presentation.Rmd
Removed warnings from lasso and KNN, so they do not show in the presentation.html

@@
 
-+ ST_Slope:
++ ST_Slope during exercise:
 
@@
 
-```{r}
+```{r message=FALSE, warning=FALSE,echo=FALSE}
 
@@
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 
@@
 # MSE values to use:
-best_lambda_lasso
-lasso_MSE
-num_zeros
+paste("Tuning Lambda",best_lambda_lasso)
+#lasso_MSE
+paste("Number of zeros",num_zeros)
 
@@
 predicted_bag <- ifelse(predict_bag>0.5,1,0)
-confm_bag<-table(predict_bag,test$HeartDisease,)
+confm_bag<-table(predicted_bag,test$HeartDisease)
 
@@
 
-- Tuning Parmeters Choosen= 600 trees, shrinkage=0.01, interaction= 2.
+- Tuning Parameters: n= 600 trees, shrinkage=0.01, interaction= 2.
 
@@
 overall_boost <- error
+overall_boost
 ```
@@
 
-paste("the overall fraction of correct predictions in the test dataset is :", acc)
+paste("The overall fraction of correct predictions in the test dataset is :", acc)
 ```

### 2021-12-08 SVM.Rmd
Everything is working, presentation html file generated.

@@
 
+{r SVM}
+
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+
+SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
+
+preds = predict(SVM_radial, test, type="class")
+acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
+
+table(preds,test$HeartDisease)
+
+overall_SVM <- acc
+
+paste("The overall fraction of correct predictions in the test dataset is :", acc)
+
 We can see from the confusion matrix of our Radial SVM that there is little bias in the miss-classifications while the MSE is `r round(radial_mse, 4)`.

### 2021-12-08 presentation.Rmd
Everything is working, presentation html file generated.

@@
 
-+ ST_Slope during exercise:
++ ST_Slope (Slope of the peak exercise ST segment):
 
@@
 #lasso_MSE
-paste("Number of zeros",num_zeros)
+#paste("Number of zeros",num_zeros)
 
@@
 
-```{r SVM}
+```{r Data, include=FALSE}
+heart_train = read.csv("train.csv")
 
-train <- read.csv("train.csv")
-test <- read.csv("test.csv")
+heart_train$Sex = factor(heart_train$Sex)
+heart_train$ChestPainType = factor(heart_train$ChestPainType)
+heart_train$FastingBS = factor(heart_train$FastingBS)
+heart_train$RestingECG = factor(heart_train$RestingECG)
+heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
+heart_train$ST_Slope = factor(heart_train$ST_Slope)
+heart_train$HeartDisease = factor(heart_train$HeartDisease)
+```
 
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
+```{r SVM_Linear}
+# Train a Linear SVM and find the best cost value
+# We set scale=TRUE becuase the continuous varibles seem to be normally distrabuted (from EDA)
+svm_linear_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="linear", data=heart_train,
+                      ranges = list(cost = seq(0.1, 5, by = 0.1)))
+svm_linear = svm_linear_cv$best.model
 
-test$ChestPainType <- as.factor(test$ChestPainType)
-test$Sex <- as.factor(test$Sex)
-test$FastingBS <- as.factor(test$FastingBS)
-test$RestingECG <- as.factor(test$RestingECG)
-test$ExerciseAngina <- as.factor(test$ExerciseAngina)
-test$ST_Slope <- as.factor(test$ST_Slope)
+preds = predict(svm_linear, heart_train, type="class")
 
-SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
+```
 
-preds = predict(SVM_radial, test, type="class")
-acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
+```{r SVM_Polynomial}
+# Train a Polynomial SVM
+# Find the best cost and degree parameter
+svm_poly_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="polynomial", data=heart_train,
+                      ranges = list(cost = seq(0.1, 5, by = 0.1),
+                                    degree = c(2,3,4, 5)))
+svm_poly = svm_poly_cv$best.model
 
-table(preds,test$HeartDisease)
+preds = predict(svm_poly, heart_train, type="class")
 
-overall_SVM <- acc
+```
 
-paste("The overall fraction of correct predictions in the test dataset is :", acc)
+```{r SVM_Radial}
+# Train a Radial SVM
+# Find the best cost and gamma parameter
+svm_radial_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="radial", data=heart_train,
+                      ranges = list(cost = seq(0.2, 5, by = 0.2),
+                                    gamma = seq(0.25, 3, by = 0.25)))
+svm_radial = svm_radial_cv$best.model
+
+preds = predict(svm_radial, heart_train, type="class")
+
+```
+
+```{r Testing}
+heart_test = read.csv("test.csv")
+
+preds_liner = predict(svm_linear, heart_test, type="class")
+acc_liner = sum(ifelse(preds_liner == heart_test$HeartDisease, 1, 0))/nrow(heart_test)
+
+preds_poly = predict(svm_poly, heart_test, type="class")
+acc_poly = sum(ifelse(preds_poly == heart_test$HeartDisease, 1, 0))/nrow(heart_test)
+
+preds_radial = predict(svm_radial, heart_test, type="class")
+acc_radial = sum(ifelse(preds_radial == heart_test$HeartDisease, 1, 0))/nrow(heart_test) #87.2%
+
+c("Linear Svm" = acc_liner, "Polynomial Svm" = acc_poly, "Radial SVM" = acc_radial)
+```
+
+- Confusion Matrix for Radial:
+
+```{r Best_SVM}
+table(heart_test$HeartDisease, preds_radial)
+overall_SVM <- acc_radial
 ```
@@
 library(knitr)
-df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  MSE = c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
+df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  Overall_Correct_Pred.= c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
 

### 2021-12-08 presentation.Rmd
Final Changes

@@
 
-paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+paste("Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))
 
@@
 
-ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill=FALSE)+
+ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill="none")+
   labs(x="Variable Name",y="Importance")
@@
 
-paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+paste("Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))
 
@@
 
-- Confusion Matrix for Radial:
+- Confusion Matrix for Radial Kernel:
 
@@
 library(knitr)
-df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  Overall_Correct_Pred.= c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
+df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  Accuracy= c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
 

### 2021-12-12 KNNModel.Rmd
adding first draft of paper and re-arranged KNN model

--- 2021-12-08 presentation.Rmd
@@
 ---
-title: "Group 9 Presentation"
-subtitle: "Heart Disease Classification"
-author: "Henrique Magalhaes Rio , Oriana Meldrum, Jorie Alvis"
-date: "12/8/2021"
-output:
-  slidy_presentation: default
-  ioslides_presentation: default
-  beamer_presentation: default
+title: "KNN Model"
+author: "Oriana Meldurm"
+date: "12/5/2021"
+output: pdf_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-options(repos='http://cran.rstudio.org')
-have.packages <- installed.packages()
-cran.packages <- c('devtools','plotrix','randomForest','tree')
-to.install <- setdiff(cran.packages, have.packages[,1])
-if(length(to.install)>0) install.packages(to.install)
-
-library(devtools)
-if(!('reprtree' %in% installed.packages())){
-  install_github('araastat/reprtree')
-}
-for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-```{r, include=FALSE}
+## R Markdown
 
-#reproducibility
-set.seed(445)
+```{r}
 
 library(ISLR)
-library(leaps)
+library(class)
 library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(glmnet)
-library(caret)
-library(pls) 
-library(tree)
-library(randomForest)
 library(knitr)
-library(class)
 library(fastDummies)
-library(reprtree)
-library(MASS)
-library(e1071)
-library(gbm)
 
@@
 
-```
-
-## Introduction
-
-In 2019 659,041 Americans died of heart disease.
-Four out of five heart disease deaths are due to heart attacks and strokes, and one-third of these deaths occur prematurely in people under 70 years of age.
-Heart failure is a common event caused by heart disease and this dataset we used contains 11 variables that can be used to predict a possible heart disease.
-
-People with cardiovascular disease or who are at high cardiovascular risk (due to the presence of one or more risk factors) need early detection and management so machine learning techniques are of great clinical intrest for early detection.
-
-(https://www.cdc.gov/nchs/fastats/leading-causes-of-death.htm, https://www.kaggle.com/fedesoriano/heart-failure-prediction)
-
-## Data {.smaller}
-
-- There are 918 observations which we split into 60/40 for training and validation.
-
-- 12 predictors are included in the data which are:
-
-+ Age
-
-+ Sex M/F
-
-+ Chest Pain Type:
-
-++ TA: Typical Angina
-
-++ ATA: Atypical Angina
-
-++ NAP: Non-Anginal Pain
-
-++ ASY: Asymptomatic
-
-+ Resting Blood Pressure [mm Hg]
-
-+ Cholesterol [mm/dl]
-
-+ Fasting Blood Sugar Y/N
-
-## Data {.smaller}
-
-+ Resting Electrocardiogram
-
-++ Normal: Normal
-
-++ ST: Having ST-T Wave abnormality
-
-++ LVH: Showing probable or definite left ventricular hypertrophy.
-
-+ Max HR: Maximum heart rate achieved.
-
-+ Exercise Angina: exercise-induced angine Y/N
-
-+ Old peak: Oldpeak= ST value of depression
-
-+ ST_Slope during exercise:
-
-++ Up sloping(up)
-
-++ Flat(flat)
-
-++ Down sloping (down)
-
-+ Heart Disease: [1:Heart Disease, 0: Normal]
-
-## Logistic Regression
-
-```{r}
-model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
-summary(model_glm)
-
-predicted2 <- predict(model_glm, newdata = test, "response")
-predicted2 <- ifelse(predicted2>0.5,1,0)
-confm2<-table(predicted=predicted2,True=test$HeartDisease)
-kable(confm2)
-
-paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-
-overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
-
-```
-
-## LDA
-
-```{r LDA}
-LDA_model = lda(HeartDisease~., data=train)
-
-preds = predict(LDA_model, test)
-
-acc = sum(ifelse(preds$class == test$HeartDisease, 1, 0))/nrow(test)
-
-table("Preds"=preds$class, "True"=test$HeartDisease)
-
-overall_LDA <- acc
-overall_LDA
-
-LDA_model
-```
-## KNN
-
-```{r message=FALSE, warning=FALSE,echo=FALSE}
-
 temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
@@
 temp_test <-temp_test[, !(names(temp_test) %in% drops)]
+
+```
+
+```{r}
+#Select best number for k
+
+set.seed(445)
+library(caret)
+
+#5 fold
+trControl <- trainControl(method = "cv", number = 5)
+
+fit <- train(as.factor(HeartDisease) ~.,
+             method = "knn", 
+             tuneGrid = expand.grid(k = 1:100), 
+             trControl = trControl, 
+             metric = "Accuracy", 
+             data = temp_train)
+
+fit
+```
+
+The best accuracy comes from a k of 23 with and accuracy of 0.6878024.
+Technically both k=1 and k=4 have higher accuracy, but we are avoiding using as any outlines would create a skew for k=1 and in certain types of classification problems, our research shower it was better to avoid using an odd number as this can tie votes.
+
+```{r}
 
@@
 
-knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)
+knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=23)
 
@@
 mean(knn_pred_1 !=test$HeartDisease)
-```
-
-## LASSO
-
-```{r message=FALSE, warning=FALSE}
-
-train_matrix <- model.matrix(HeartDisease ~ ., data = train)
-test_matrix <- model.matrix(HeartDisease ~ ., data = test)
-grid <- 10^seq(10, -2, length=100)
-
-lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
-#plot(lasso)
-
-cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
-#plot(cv_lasso)
-best_lambda_lasso <- cv_lasso$lambda.min
-
-lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
-lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
-
-num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
-
-# MSE values to use:
-paste("Tuning Lambda",best_lambda_lasso)
-#lasso_MSE
-paste("Number of zeros",num_zeros)
-
-#Confusion Matrix
-predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
-predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
-confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
-confm_lasso
-
-#Over all correct:
-overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
-
-kable(confm_lasso) #confusion matrix
-overall_lasso #over all correct
-```
-
-## Bagging
-```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
-train <- read.csv("train.csv")
-test <- read.csv("test.csv")
-
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
-
-test$ChestPainType <- as.factor(test$ChestPainType)
-test$Sex <- as.factor(test$Sex)
-test$FastingBS <- as.factor(test$FastingBS)
-test$RestingECG <- as.factor(test$RestingECG)
-test$ExerciseAngina <- as.factor(test$ExerciseAngina)
-test$ST_Slope <- as.factor(test$ST_Slope)
-
-bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
-                        importance = TRUE,
-                        proximity = TRUE)
-
-#Confusion Matrix
-predict_bag <- predict(bagging, newdata = test)
-predicted_bag <- ifelse(predict_bag>0.5,1,0)
-confm_bag<-table(predicted_bag,test$HeartDisease)
-
-#MSE
-bagging_mse <- mean((predict_bag - test$HeartDisease)^2)
-
-#Over all correct:
-overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
-
-kable(confm_bag) #confusion matrix
-overall_bag #over all correct
-
-#Visulizaton of bagging
-VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
-
-#sort variable importance descending
-#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
-
-#visualize variable importance with horizontal bar plot
-#barplot(VI_plot$imp.0,
-   #     names.arg=rownames(VI_plot),
-    #    horiz=TRUE,
-     #   col='steelblue',
-      #  xlab='Variable Importance')
-
-ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill=FALSE)+
-  labs(x="Variable Name",y="Importance")
-```
-
-## Random Forest
-
-```{r}
-rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)-1), importance = TRUE, type= "classification")
-
-true <- test$HeartDisease
-rf <- predict(rf_fit,test,type="class")
-
-confm2<-table(predicted=rf,True=true)
-confm2
-
-paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-
-overall_RF <- (confm2[1,1]+confm2[2,2])/sum(confm2)
-rf_fit
-
-imp<-data.frame(rf_fit$importance)
-
-imp$variables <- rownames(imp)
-rownames(imp) <- 1:nrow(imp)
-
-imp <- imp[order(imp$MeanDecreaseAccuracy),]
 
 ```
-
-## Random Forest Plot
-
-```{r}
-ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
-```
-
-## Representative Tree
-- Following the paper "Identifying representative Trees From Ensembles"  by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
-
-- This tree is calculated using a measure of similarity, in which 2 trees are similar if they have the same prediction for all observations.
-
-- The representative tree is then choosen based on its averaged similarity between all other trees.
-
-- Using this idea on the random forest we get the following tree:
-
-## Representative Tree Plot
-```{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
-reptree <- ReprTree(rf_fit, train, metric='d2')
-
-plot(reptree, index=1)
-
-```
-
-## Boosting
-
-```{r}
-
-train <- read.csv("train.csv")
-test <- read.csv("test.csv")
-
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
-
-test$ChestPainType <- as.factor(test$ChestPainType)
-test$Sex <- as.factor(test$Sex)
-test$FastingBS <- as.factor(test$FastingBS)
-test$RestingECG <- as.factor(test$RestingECG)
-test$ExerciseAngina <- as.factor(test$ExerciseAngina)
-test$ST_Slope <- as.factor(test$ST_Slope)
-
-set.seed(445)
-
-oob = trainControl(method = "oob")
-cv_5 = trainControl(method = "cv", number = 5)
-
-gbm_grid =  expand.grid(interaction.depth = 1:5,
-                        n.trees = (1:5) * 200,
-                        shrinkage = c(0.01,0.015,0.1),
-                        n.minobsinnode = 10)
-
-gbm_tune = train(as.factor(HeartDisease) ~ ., data = train,
-                      method = "gbm",
-                      trControl = cv_5,
-                      verbose = FALSE,
-                      tuneGrid = gbm_grid)
-
-plot(gbm_tune)
-
-```
-
-- Tuning Parameters: n= 600 trees, shrinkage=0.01, interaction= 2.
-
-```{r}
-boost <- gbm(HeartDisease~.,data=train,n.trees=600,shrinkage=0.01,interaction.depth = 2 ,distribution = "bernoulli")
-
-boostpred = ifelse(predict(boost,test, n.trees = 500, "response") > 0.5, "1", "0")
-
-confboost<-table(predicted = boostpred, actual = test$HeartDisease)
-confboost
-error<-(confboost[1,1]+confboost[2,2])/sum(confboost)
-overall_boost <- error
-overall_boost
-```
-
-```{r,include=FALSE}
-summa<-summary(boost)
-```
-
-```{r}
-ggplot(summa)+geom_col(aes(x=var,y=rel.inf,fill=var))+coord_flip()+labs(y="Relative Influence", x= " Variable")+guides(fill="none")
-```
-
-## SVM
-
-```{r SVM}
-
-train <- read.csv("train.csv")
-test <- read.csv("test.csv")
-
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
-
-test$ChestPainType <- as.factor(test$ChestPainType)
-test$Sex <- as.factor(test$Sex)
-test$FastingBS <- as.factor(test$FastingBS)
-test$RestingECG <- as.factor(test$RestingECG)
-test$ExerciseAngina <- as.factor(test$ExerciseAngina)
-test$ST_Slope <- as.factor(test$ST_Slope)
-
-SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
-
-preds = predict(SVM_radial, test, type="class")
-acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
-
-table(preds,test$HeartDisease)
-
-overall_SVM <- acc
-
-paste("The overall fraction of correct predictions in the test dataset is :", acc)
-```
-
-## Results
-
-- What method would be the best?
-
-```{r}
-library(knitr)
-df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  MSE = c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
-
-kable(df1)
-
-```

### 2021-12-12 paper.Rmd
adding first draft of paper and re-arranged KNN model

---
title: "Group 9 Paper"
author: "Henrique Magalhaes Rio , Oriana Meldrum, Jorie Alvis"
date: "12/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In 2019 659,041 Americans died of heart disease. Four out of five heart disease deaths are due to heart attacks and strokes, and one-third of these deaths occur prematurely in people under 70 years of age. Heart failure is a common event caused by heart disease and this dataset we used contains 11 variables that can be used to predict a possible heart disease.

People with cardiovascular disease or who are at high cardiovascular risk (due to the presence of one or more risk factors) need early detection and management so machine learning techniques are of great clinical interest for early detection.

# Methods

## Logistic Regression

This will be the section on Logistic Regression. Include discussion on how & why you chose to do it the way you did it.

## LDA

This will be the section on LDA. Include discussion on how & why you chose to do it the way you did it.

## KNN

K-nearest Neighbors is useful for both regression and classification. An object will be classified by a vote of the nearest k neighbors. The object is assigned to the class most common to those k neighbors. For this project, k was selected by cross validation.  The best accuracy comes from a k of 23 with and accuracy of 0.6878024. Technically both k=1 and k=4 have higher accuracy, but we are avoiding k=1 using as any outlines would create a skew and in certain types of classification problems, our research shower it was better to avoid using an odd number as this can tie votes. With an prediction rate of only 0.6757493 had the lowest overall correct prediction rate.

## LASSO

When using the LASSO method, it was important to select family = "binomial" for the classification problem. The tuning lambda value was selected through cross validation.

## Bagging

This will be the section on Bagging. Include discussion on how & why you chose to do it the way you did it.

## Bosting

This will be the section on Boosting. Include discussion on how & why you chose to do it the way you did it.

## Random Forest

This will be the section on Random Forest. Include discussion on how & why you chose to do it the way you did it.

## Representative Tree

This will be the section on Representative Tree. Include discussion on how & why you chose to do it the way you did it.

## SVM

This will be the section on SVM. Include discussion on how & why you chose to do it the way you did it.

# Methods That Didn't Work

This will be the section on methods we considered but did work. Include discussion on how & why you didn't chose to do it.

# Conclusion

## Future Proposal

It would be interesting to re-try this project, removing the ST-Slope value. As we learned from industry professionals, the ST Slope is nearly always associated with some type of heart rhythm anomaly, and thus is frequently present in cases of heart disease. Many of our models indicated the ST Slope as the number one predictor for heart disease due to this strong association.

### 2021-12-12 paper.Rmd
adding bagging dicussion

@@
 
-This will be the section on Bagging. Include discussion on how & why you chose to do it the way you did it.
+The bagging model was preformed with 500 trees and identified the most important variable as being ST_Slope followed by Chest Pain Type.
+Interestingly, resting ECG and Resting BP both had negative importance.
+This means if we remove those two features, we would improve model performance.
+As other methods found resting ECG and Resting BP to be among the least important variable, it would be worth taking the time to remove them and see if it improves the accuracy rates for our data.
 

### 2021-12-13 paper.Rmd
Added Logistic regression code and write up

--- 2021-12-12 paper.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
+
+set.seed(445)
+
+library(ISLR)
+library(leaps)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(glmnet)
+library(caret)
+library(pls) 
+library(tree)
+library(randomForest)
+library(knitr)
+library(class)
+library(fastDummies)
+library(reprtree)
+library(MASS)
+library(e1071)
+library(gbm)
+
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+train$HeartDisease <- as.factor(train$HeartDisease)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+test$HeartDisease <- as.factor(test$HeartDisease)
 ```
@@
 
-This will be the section on Logistic Regression. Include discussion on how & why you chose to do it the way you did it.
+```{r,echo=FALSE}
+model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
+summary(model_glm)
+
+predicted2 <- predict(model_glm, newdata = test, "response")
+predicted2 <- ifelse(predicted2>0.5,1,0)
+confm2<-table(predicted=predicted2,True=test$HeartDisease)
+confm2
+
+paste("Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+
+overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
+```
+
+ We started by fitting a logistic regression model with all the variables, and from the results above we see that there are a few non-significant variables and important ones, it is worth noting that SexM variable is is statistically significant and positive, which interesting as it means that on average men are more likely to get heart disease than women given that the other variables are the same, also, ST_SlopeFlat and ST_SlopeUp are statistically significant and the  show different signs which means that people with ST_SlopeFlat have a higher Likelyhood of having heart Disease when compared to those that have ST_SlopeUp, and this is consistent with science as ST_Slope is a important factor in the diagnosis of a heart Disease.
 

### 2021-12-13 paper.Rmd
Added Code for boosting CV

@@
 
- We started by fitting a logistic regression model with all the variables, and from the results above we see that there are a few non-significant variables and important ones, it is worth noting that SexM variable is is statistically significant and positive, which interesting as it means that on average men are more likely to get heart disease than women given that the other variables are the same, also, ST_SlopeFlat and ST_SlopeUp are statistically significant and the  show different signs which means that people with ST_SlopeFlat have a higher Likelyhood of having heart Disease when compared to those that have ST_SlopeUp, and this is consistent with science as ST_Slope is a important factor in the diagnosis of a heart Disease.
+ We started by fitting a logistic regression model with all the variables, and from the results above we see that there are a few non-significant variables and important ones, it is worth noting that SexM variable is is statistically significant and positive, which interesting as it means that on average men are more likely to get heart disease than women given that the other variables are the same, also, ST_SlopeFlat and ST_SlopeUp are statistically significant and the  show different signs which means that people with ST_SlopeFlat have a higher Likelihood of having heart Disease when compared to those that have ST_SlopeUp, and this is consistent with science as ST_Slope is a important factor in the diagnosis of a heart Disease.
+Overall, for a simple method that also allows for inference, it performed relatively well with an accuracy of 0.85.
 
@@
 ## Bosting
+
+```{r}
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+
+set.seed(445)
+
+oob = trainControl(method = "oob")
+cv_5 = trainControl(method = "cv", number = 5)
+
+gbm_grid =  expand.grid(interaction.depth = 1:5,
+                        n.trees = (1:5) * 200,
+                        shrinkage = c(0.01,0.015,0.1),
+                        n.minobsinnode = 10)
+
+gbm_tune = train(as.factor(HeartDisease) ~ ., data = train,
+                      method = "gbm",
+                      trControl = cv_5,
+                      verbose = FALSE,
+                      tuneGrid = gbm_grid)
+
+plot(gbm_tune)
+```
 

### 2021-12-13 paper.Rmd
Boosting fit and write up

@@
 
-```{r}
+```{r,echo=FALSE}
 train <- read.csv("train.csv")
@@
                         n.trees = (1:5) * 200,
-                        shrinkage = c(0.01,0.015,0.1),
+                        shrinkage = c(0.001,0.01,0.015,0.1),
                         n.minobsinnode = 10)
@@
                       tuneGrid = gbm_grid)
+plot(gbm_tune)
 
-plot(gbm_tune)
+gbm_tune$bestTune
+
 ```
 
-This will be the section on Boosting. Include discussion on how & why you chose to do it the way you did it.
+In order to tune tree boosting method I used a 5 fold cross-validation on the training data set, from the plot it seems that there is a spike on the on the accuracy when using shrinkage with 0.01, 400 trees and a interaction depth of 2.
+It should also be noted that several other ranges of the tune parameters were tested but for simplicity sake on the range on the plot was left which included the best one overall.
+
+```{r,echo=FALSE}
+boost <- gbm(HeartDisease~.,data=train,n.trees=400,shrinkage=0.01,interaction.depth = 2 ,distribution = "bernoulli")
+
+boostpred = ifelse(predict(boost,test, n.trees = 400, "response") > 0.5, "1", "0")
+
+confboost<-table(predicted = boostpred, actual = test$HeartDisease)
+confboost
+error<-(confboost[1,1]+confboost[2,2])/sum(confboost)
+overall_boost <- error
+paste("accuraccy:",overall_boost)
+```
+Fitting the model with the tuning parameters previously discussed, we get a much better accuracy when compared to the other methods but at the cost of having lost the ability of doing inference in this data set.
+We can take a look at the importance graph below, in which we can see that ST_Slope, ChestPainType, and Cholesterol are the most important variables, which is somewhat consistent with most of the models used so far.
+Also Consistent with the other models, we can see that RestingECG and RestingBP are the least important variables for boosting.
+
+```{r,include=FALSE}
+summa<-summary(boost)
+```
+
+```{r,echo=FALSE}
+ggplot(summa)+geom_col(aes(x=var,y=rel.inf,fill=var))+coord_flip()+labs(y="Relative Influence", x= " Variable")+guides(fill="none")
+```
 

### 2021-12-16 paper.Rmd
added code for rf

@@
 
-This will be the section on Random Forest. Include discussion on how & why you chose to do it the way you did it.
+```{r}
+rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)-1), importance = TRUE, type= "classification", ntrees=1000)
+
+true <- test$HeartDisease
+rf <- predict(rf_fit,test,type="class")
+
+confm2<-table(predicted=rf,True=true)
+confm2
+
+paste(" Test Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))
+
+overall_RF <- (confm2[1,1]+confm2[2,2])/sum(confm2)
+rf_fit
+
+imp<-data.frame(rf_fit$importance)
+
+imp$variables <- rownames(imp)
+rownames(imp) <- 1:nrow(imp)
+
+imp <- imp[order(imp$MeanDecreaseAccuracy),]
+```
 

### 2021-12-16 paper.Rmd
RF written part

@@
 ```{r}
-rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)-1), importance = TRUE, type= "classification", ntrees=1000)
+rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)), importance = TRUE, type= "classification", ntrees=1000)
 
@@
 
+Next, we fitted a Random Forest, in this case I chose the number of predictors randomly sampled at each iteration to be $m=\sqrt{p}$, as this was what was recommended by the class book.
+For random forests, the test accuracy relatively good at 0.8664 which is good, but not better than some of te other methods.
+Looking at the variable importance, we have some similar results to the other tree based methods, as ST_Slope, ChestPainType, and Cholesterol seem to be the most important variables.
+However, in random forests we got a negative value for the RestingBP, which is somewhat consistent with the other methos as RestingBP also did not seem to be a important variable in dertermining heart disease, however, in this case it means that for random forests is seems that including RestingBP decreases the accuracy of the method.
+
+```{r}
+ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
+```
+
 ## Representative Tree
-
-This will be the section on Representative Tree. Include discussion on how & why you chose to do it the way you did it.
 

### 2021-12-16 paper.Rmd
Representatie tree write up

@@
 
-```{r}
+```{r,echo=FALSE}
 rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)), importance = TRUE, type= "classification", ntrees=1000)
@@
 
-```{r}
+```{r,echo=FALSE}
 ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
@@
 ## Representative Tree
+
+While working on the random forests, I began to  search for a way to visualize at least one of the trees in the ensemble, in order to more visualization and perhaps get a deeper insight into the method.
+While I found a way to do it in stack overflow, I also found a recommendation of a paper by Banerjee, et al.
+called "Identifying representative trees from ensembles" (2012), which was published in the journal of statistical medicine.
+What we found really interesting is that the example application which was very similar to our application as it was classification problem using kidney cancer data, and there they discuss the importance of not losing the ability to do inference while still having high accuracy.
+
+In the paper Banerjee, et al. discuss how to choose the most representative tree in a n esemble.
 

### 2021-12-16 paper.Rmd
more rep tree

@@
 
-In the paper Banerjee, et al. discuss how to choose the most representative tree in a n esemble.
+In the paper Banerjee, et al.
+discuss how to choose the most representative tree in an ensemble, which is done by finding the average distance between one tree and all other trees in the ensemble.
+Banerjee, et al.
+discusses several types of distance in this paper only one of them was used, as it was already implemented in R package.
+This measure captures the similarities between the predictions, in that 2 trees are similar if they have the same predictions outcomes
 

### 2021-12-16 paper.Rmd
Adding final edits to O's methods

--- 2021-12-16 paper.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-
-set.seed(445)
-
-library(ISLR)
-library(leaps)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(glmnet)
-library(caret)
-library(pls) 
-library(tree)
-library(randomForest)
-library(knitr)
-library(class)
-library(fastDummies)
-library(reprtree)
-library(MASS)
-library(e1071)
-library(gbm)
-
-train <- read.csv("train.csv")
-test <- read.csv("test.csv")
-
-train$ChestPainType <- as.factor(train$ChestPainType)
-train$Sex <- as.factor(train$Sex)
-train$FastingBS <- as.factor(train$FastingBS)
-train$RestingECG <- as.factor(train$RestingECG)
-train$ExerciseAngina <- as.factor(train$ExerciseAngina)
-train$ST_Slope <- as.factor(train$ST_Slope)
-train$HeartDisease <- as.factor(train$HeartDisease)
-
-test$ChestPainType <- as.factor(test$ChestPainType)
-test$Sex <- as.factor(test$Sex)
-test$FastingBS <- as.factor(test$FastingBS)
-test$RestingECG <- as.factor(test$RestingECG)
-test$ExerciseAngina <- as.factor(test$ExerciseAngina)
-test$ST_Slope <- as.factor(test$ST_Slope)
-test$HeartDisease <- as.factor(test$HeartDisease)
 ```
@@
 
+## Motivation
+
 People with cardiovascular disease or who are at high cardiovascular risk (due to the presence of one or more risk factors) need early detection and management so machine learning techniques are of great clinical interest for early detection.
+
+For this project, we proposed to use the various machine learning methods learned this semester in CSU's DSCI 445 Machine Learning class that can be applied to a classification problem.
 
@@
 
-```{r,echo=FALSE}
-model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
-summary(model_glm)
-
-predicted2 <- predict(model_glm, newdata = test, "response")
-predicted2 <- ifelse(predicted2>0.5,1,0)
-confm2<-table(predicted=predicted2,True=test$HeartDisease)
-confm2
-
-paste("Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-
-overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
-```
-
- We started by fitting a logistic regression model with all the variables, and from the results above we see that there are a few non-significant variables and important ones, it is worth noting that SexM variable is is statistically significant and positive, which interesting as it means that on average men are more likely to get heart disease than women given that the other variables are the same, also, ST_SlopeFlat and ST_SlopeUp are statistically significant and the  show different signs which means that people with ST_SlopeFlat have a higher Likelihood of having heart Disease when compared to those that have ST_SlopeUp, and this is consistent with science as ST_Slope is a important factor in the diagnosis of a heart Disease.
-Overall, for a simple method that also allows for inference, it performed relatively well with an accuracy of 0.85.
+This will be the section on Logistic Regression. Include discussion on how & why you chose to do it the way you did it.
 
@@
 
+The next method we used was K-Nearest Neighbors.
 K-nearest Neighbors is useful for both regression and classification.
@@
 Technically both k=1 and k=4 have higher accuracy, but we are avoiding k=1 using as any outlines would create a skew and in certain types of classification problems, our research shower it was better to avoid using an odd number as this can tie votes.
-With an prediction rate of only 0.6757493 had the lowest overall correct prediction rate.
+With an accuracy rate of only 0.6757493 had the lowest overall accuracy rate.
+
+```{r message=FALSE, warning=FALSE,echo=FALSE}
+temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
+temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))
+drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
+temp_train <- temp_train[, !(names(temp_train) %in% drops)]
+temp_test <-temp_test[, !(names(temp_test) %in% drops)]
+#trainknn <- temp_train[-temp_train$HeartDisease]
+#testknn <- test[-test$HeartDisease]
+num_k <- seq(1:100)
+my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
+knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)
+confm_knn_1<-table(knn_pred_1,test$HeartDisease)
+overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)
+overall_knn_1
+ #confusion matrix
+kable(confm_knn_1)
+mean(knn_pred_1 !=test$HeartDisease)
+
+```
 
@@
 
-When using the LASSO method, it was important to select family = "binomial" for the classification problem. The tuning lambda value was selected through cross validation.
+After KNN, we used  the least absolute shrinkage and selection operator(LASSO).
+When using the LASSO method, it was important to select family = "binomial" for the classification problem.
+The tuning lambda value was selected through cross validation.
+LASSO performed relatively well with an accuracy of 0.85.
+
+```{r message=FALSE, warning=FALSE}
+train_matrix <- model.matrix(HeartDisease ~ ., data = train)
+test_matrix <- model.matrix(HeartDisease ~ ., data = test)
+grid <- 10^seq(10, -2, length=100)
+lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
+#plot(lasso)
+cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
+#plot(cv_lasso)
+best_lambda_lasso <- cv_lasso$lambda.min
+lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
+lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
+num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
+# MSE values to use:
+paste("Tuning Lambda",best_lambda_lasso)
+#lasso_MSE
+#paste("Number of zeros",num_zeros)
+#Confusion Matrix
+predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
+predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
+confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
+confm_lasso
+#Over all correct:
+overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
+kable(confm_lasso) #confusion matrix
+overall_lasso #over all correct
+```
 
@@
 
-This will be the section on Bagging. Include discussion on how & why you chose to do it the way you did it.
+The bagging model was preformed with 500 trees and identified the most important variable as being ST_Slope followed by Chest Pain Type.
+Interestingly, resting ECG and Resting BP both had negative importance.
+This means if we remove those two features, we would improve model performance.
+As other methods found resting ECG and Resting BP to be among the least important variable, it would be worth taking the time to remove them and see if it improves the accuracy rates for our data.
+Bagging performed rather disappointingly with an accuracy of 0.84, the lowest of all methods except KNN.
 
-## Bosting
-
-```{r,echo=FALSE}
+```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
 train <- read.csv("train.csv")
 test <- read.csv("test.csv")
-
 train$ChestPainType <- as.factor(train$ChestPainType)
@@
 train$ST_Slope <- as.factor(train$ST_Slope)
-
 test$ChestPainType <- as.factor(test$ChestPainType)
@@
 test$ST_Slope <- as.factor(test$ST_Slope)
-
-set.seed(445)
-
-oob = trainControl(method = "oob")
-cv_5 = trainControl(method = "cv", number = 5)
-
-gbm_grid =  expand.grid(interaction.depth = 1:5,
-                        n.trees = (1:5) * 200,
-                        shrinkage = c(0.001,0.01,0.015,0.1),
-                        n.minobsinnode = 10)
-
-gbm_tune = train(as.factor(HeartDisease) ~ ., data = train,
-                      method = "gbm",
-                      trControl = cv_5,
-                      verbose = FALSE,
-                      tuneGrid = gbm_grid)
-plot(gbm_tune)
-
-gbm_tune$bestTune
-
+bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
+                        importance = TRUE,
+                        proximity = TRUE) 
+#Confusion Matrix
+predict_bag <- predict(bagging, newdata = test)
+predicted_bag <- ifelse(predict_bag>0.5,1,0)
+confm_bag<-table(predicted_bag,test$HeartDisease)
+#MSE
+bagging_mse <- mean((predict_bag - test$HeartDisease)^2) 
+#Over all correct:
+overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
+kable(confm_bag) #confusion matrix
+overall_bag #over all correct
+#Visulizaton of bagging
+VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
+#sort variable importance descending
+#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
+#visualize variable importance with horizontal bar plot
+#barplot(VI_plot$imp.0,
+   #     names.arg=rownames(VI_plot),
+    #    horiz=TRUE,
+     #   col='steelblue',
+      #  xlab='Variable Importance')
+ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill="none")+
+  labs(x="Variable Name",y="Importance")
 ```
 
-In order to tune tree boosting method I used a 5 fold cross-validation on the training data set, from the plot it seems that there is a spike on the on the accuracy when using shrinkage with 0.01, 400 trees and a interaction depth of 2.
-It should also be noted that several other ranges of the tune parameters were tested but for simplicity sake on the range on the plot was left which included the best one overall.
+## Bosting
 
-```{r,echo=FALSE}
-boost <- gbm(HeartDisease~.,data=train,n.trees=400,shrinkage=0.01,interaction.depth = 2 ,distribution = "bernoulli")
-
-boostpred = ifelse(predict(boost,test, n.trees = 400, "response") > 0.5, "1", "0")
-
-confboost<-table(predicted = boostpred, actual = test$HeartDisease)
-confboost
-error<-(confboost[1,1]+confboost[2,2])/sum(confboost)
-overall_boost <- error
-paste("accuraccy:",overall_boost)
-```
-Fitting the model with the tuning parameters previously discussed, we get a much better accuracy when compared to the other methods but at the cost of having lost the ability of doing inference in this data set.
-We can take a look at the importance graph below, in which we can see that ST_Slope, ChestPainType, and Cholesterol are the most important variables, which is somewhat consistent with most of the models used so far.
-Also Consistent with the other models, we can see that RestingECG and RestingBP are the least important variables for boosting.
-
-```{r,include=FALSE}
-summa<-summary(boost)
-```
-
-```{r,echo=FALSE}
-ggplot(summa)+geom_col(aes(x=var,y=rel.inf,fill=var))+coord_flip()+labs(y="Relative Influence", x= " Variable")+guides(fill="none")
-```
+This will be the section on Boosting. Include discussion on how & why you chose to do it the way you did it.
 
@@
 
-```{r}
-rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)), importance = TRUE, type= "classification", ntrees=1000)
-
-true <- test$HeartDisease
-rf <- predict(rf_fit,test,type="class")
-
-confm2<-table(predicted=rf,True=true)
-confm2
-
-paste(" Test Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))
-
-overall_RF <- (confm2[1,1]+confm2[2,2])/sum(confm2)
-rf_fit
-
-imp<-data.frame(rf_fit$importance)
-
-imp$variables <- rownames(imp)
-rownames(imp) <- 1:nrow(imp)
-
-imp <- imp[order(imp$MeanDecreaseAccuracy),]
-```
-
-Next, we fitted a Random Forest, in this case I chose the number of predictors randomly sampled at each iteration to be $m=\sqrt{p}$, as this was what was recommended by the class book.
-For random forests, the test accuracy relatively good at 0.8664 which is good, but not better than some of te other methods.
-Looking at the variable importance, we have some similar results to the other tree based methods, as ST_Slope, ChestPainType, and Cholesterol seem to be the most important variables.
-However, in random forests we got a negative value for the RestingBP, which is somewhat consistent with the other methos as RestingBP also did not seem to be a important variable in dertermining heart disease, however, in this case it means that for random forests is seems that including RestingBP decreases the accuracy of the method.
-
-```{r}
-ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
-```
+This will be the section on Random Forest. Include discussion on how & why you chose to do it the way you did it.
 
 ## Representative Tree
+
+This will be the section on Representative Tree. Include discussion on how & why you chose to do it the way you did it.
 

### 2021-12-16 paper.Rmd

--- 2021-12-16 paper.Rmd
@@
 
-This will be the section on Bagging. Include discussion on how & why you chose to do it the way you did it.
+The bagging model was preformed with 500 trees and identified the most important variable as being ST_Slope followed by Chest Pain Type.
+Interestingly, resting ECG and Resting BP both had negative importance.
+This means if we remove those two features, we would improve model performance.
+As other methods found resting ECG and Resting BP to be among the least important variable, it would be worth taking the time to remove them and see if it improves the accuracy rates for our data.
+Bagging performed rather disappointingly with an accuracy of 0.84, the lowest of all methods except KNN.
+
+```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
+                        importance = TRUE,
+                        proximity = TRUE) 
+#Confusion Matrix
+predict_bag <- predict(bagging, newdata = test)
+predicted_bag <- ifelse(predict_bag>0.5,1,0)
+confm_bag<-table(predicted_bag,test$HeartDisease)
+#MSE
+bagging_mse <- mean((predict_bag - test$HeartDisease)^2) 
+#Over all correct:
+overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
+kable(confm_bag) #confusion matrix
+overall_bag #over all correct
+#Visulizaton of bagging
+VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
+#sort variable importance descending
+#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
+#visualize variable importance with horizontal bar plot
+#barplot(VI_plot$imp.0,
+   #     names.arg=rownames(VI_plot),
+    #    horiz=TRUE,
+     #   col='steelblue',
+      #  xlab='Variable Importance')
+ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill="none")+
+  labs(x="Variable Name",y="Importance")
+```
 

### 2021-12-16 paper.Rmd
repr tree
Reprtree

@@
 Banerjee, et al.
-discusses several types of distance in this paper only one of them was used, as it was already implemented in R package.
-This measure captures the similarities between the predictions, in that 2 trees are similar if they have the same predictions outcomes
+discusses several types of distance in this paper only one of them was used, as it was already implemented in a R package.
+This measure captures the similarities between the predictions, in that 2 trees are similar if they have the same predictions for all subjects, this is defined by Banerjee, et al as :
+$d_2(T_1,T_2)=\frac{1}{n}\sum_{i=1}^n(\hat{y}_{1i}-\hat{y}_{2i})^2$, if the trees perfectly similar, the distance $d_2(T_1,T_2)$ is equal 0.
+For for finding the representative tree we used the package "reptree", and using it on this random forests we got the following tree:
 

### 2021-12-16 paper.Rmd
added reprtree code without output, still missing installation for now.

@@
 ```{r,echo=FALSE}
+set.seed(445)
+
 rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)), importance = TRUE, type= "classification", ntrees=1000)
@@
 Next, we fitted a Random Forest, in this case I chose the number of predictors randomly sampled at each iteration to be $m=\sqrt{p}$, as this was what was recommended by the class book.
-For random forests, the test accuracy relatively good at 0.8664 which is good, but not better than some of te other methods.
+For random forests, the test accuracy relatively good at 0.8719 which is good, but not better than some of te other methods.
 Looking at the variable importance, we have some similar results to the other tree based methods, as ST_Slope, ChestPainType, and Cholesterol seem to be the most important variables.
@@
 
+{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
+set.seed(445)
+
+reptree <- ReprTree(rf_fit, train, metric='d2')
+
+plot(reptree, index=1)
+
+In the tree we can see ST_Slope is the root node, which was expected since most models considered it the most important variable, however, it is interesting that in the next nodes we have FastingBS and Sex:F both variables were not considered very important by the random forest model.
+This made us somewhat skeptical about this algorithm, however, it does make sense as we are looking for a representative tree not the most accuracate tree, if we had more time we would've liked to implement our own algorithm based of the paper in order to further validate the results.
+
 ## SVM

### 2021-12-16 paper.Rmd
Update paper.Rmd

--- 2021-12-16 paper.Rmd
@@
 
-This will be the section on LDA. Include discussion on how & why you chose to do it the way you did it.
+Our next method for our binary classification problem was Linear Discriminate Analysis (LDA).
+LDA is used to find a linear combination of factors to sperate 2 or more classes.
+Typically, LDA uses only numeric predictors but through dummy variables, LDA can use categorical variables.
+This allows for limited inferences to be made from LDA.
+
+```{r LDA}
+LDA_model = lda(HeartDisease~., data=train)
+LDA_model
+
+preds = predict(LDA_model, test)
+overall_LDA = sum(ifelse(preds$class == test$HeartDisease, 1, 0))/nrow(test)
+
+table(preds$class, "True"=heart_test$HeartDisease)
+```
+
+The LDA model achived a test accuracy of `r round(overall_LDA, 3)`.
+Even though we can't make strong infrences from our LDA, we can look at the means of each factor of each class.
+We see that the Resting BP factor has little class mean difference.
 
@@
 
-This will be the section on SVM. Include discussion on how & why you chose to do it the way you did it.
+The last model we used was SVM.
+We wanted to using the tuning method to best optimize our SVM.
+We tested three kernals for our SVM: linear, polynomial, and radial.
+We tuned the cost parameter for each, the degree parameter for the polynomial kernal, and the gamma parameter for the radial kernal.
+
+```{r SVM_linear}
+svm_linear_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="linear", data=train,
+                      ranges = list(cost = seq(0.1, 5, by = 0.1)))
+
+ggplot(data=svm_linear_cv$performances) + geom_point(aes(x=cost, y=error))
+```
+
+```{r SVM_poly}
+svm_poly_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="polynomial", data=train,
+                      ranges = list(cost = seq(0.1, 5, by = 0.1),
+                                    degree = c(2,3,4, 5)))
+
+ggplot(data=svm_poly_cv$performances) + geom_point(aes(x=cost, y=error)) + 
+  facet_grid(~degree) + labs(title="Error of Polynomial Kernal by Degree")
+```
+
+```{r SVM_radial}
+SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
+
+preds = predict(SVM_radial, test, type="class")
+acc = sum(ifelse(preds == test$HeartDisease, 1, 0))/nrow(test)
+
+table(preds,test$HeartDisease)
+
+overall_SVM <- acc
+```
+
+We found that the radial kernal with cost of 0.4 and gamma parameter of 0.25 gave us the best training accuracy. Our testing accuracy for our SVM model was `r round(acc, 3)`.
 
@@
 
+We found that all of our models, except KNN, had an accuracy within a few points.
+It was interesting to note from LDA that our prioir probabilities of our two classes closely follow the CDCs probability of getting heart disease.
+ST-Slope was shown to be the most important predictor in most of our models and we estimate this is the reason for the similar accuracies between models.
+
 ## Future Proposal
@@
 Many of our models indicated the ST Slope as the number one predictor for heart disease due to this strong association.
+We also would've liked to be able to make more infrences from our models.
+A model like LDA may have one of the highest accuracies, but we could make very little infrences about our data from this model.

### 2021-12-16 paper.Rmd

@@
 For for finding the representative tree we used the package "reptree", and using it on this random forests we got the following tree:
+
+{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
+set.seed(445)
+
+reptree <- ReprTree(rf_fit, train, metric='d2')
+
+plot(reptree, index=1)
+
+In the tree we can see ST_Slope is the root node, which was expected since most models considered it the most important variable, however, it is interesting that in the next nodes we have FastingBS and Sex:F both variables were not considered very important by the random forest model.
+This made us somewhat skeptical about this algorithm, however, it does make sense as we are looking for a representative tree not the most accuracate tree, if we had more time we would've liked to implement our own algorithm based of the paper in order to further validate the results.
 

### 2021-12-16 paper.Rmd
fixes

@@
 
-table(preds$class, "True"=heart_test$HeartDisease)
+table(preds$class, "True"=test$HeartDisease)
 ```
@@
 
-{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
+```{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
 set.seed(445)
@@
 plot(reptree, index=1)
+
+```
 
@@
 
-table(preds,test$HeartDisease)
+#table(preds,test$HeartDisease)
 
@@
 
-We found that the radial kernal with cost of 0.4 and gamma parameter of 0.25 gave us the best training accuracy. Our testing accuracy for our SVM model was `r round(acc, 3)`.
-
-# Methods That Didn't Work
-
-This will be the section on methods we considered but did work. Include discussion on how & why you didn't chose to do it.
+We found that the radial kernel with cost of 0.4 and gamma parameter of 0.25 gave us the best training accuracy. Our testing accuracy for our SVM model was `r round(acc, 3)`.
 
@@
 
+```{r}
+library(knitr)
+df1 = data.frame(Method = c("Logistic Regression","LDA","KNN","LASSO","Bagging","Random Forest","Boosting","SVM"),  Accuracy= c(overall_logit,overall_LDA,overall_knn_1,overall_lasso,overall_bag,overall_RF,overall_boost,overall_SVM))
+
+kable(df1)
+
+```
+
 We found that all of our models, except KNN, had an accuracy within a few points.
-It was interesting to note from LDA that our prioir probabilities of our two classes closely follow the CDCs probability of getting heart disease.
-ST-Slope was shown to be the most important predictor in most of our models and we estimate this is the reason for the similar accuracies between models.
+It was interesting to note from LDA that our prior probabilities of our two classes closely follow the CDCs probability of getting heart disease.
+ST-Slope was shown to be the most important predictor in most of our models and we estimate this is the reason for the similar accuracy between models.
+So choosing the best model depends a lot on whether or not we value more inference or prediction, if we only care about inference, boosting is probably the best method otherwise, something like the logistic regression might be more useful for inference, however, if both are really important a using a random forest with a representative tree might be the best way.
 
@@
 Many of our models indicated the ST Slope as the number one predictor for heart disease due to this strong association.
-We also would've liked to be able to make more infrences from our models.
-A model like LDA may have one of the highest accuracies, but we could make very little infrences about our data from this model.
+We also would've liked to be able to make more inference from our models.
+A model like LDA may have one of the highest accuracy, but we could make very little inference about our data from this model.

### 2021-12-16 paper.Rmd
Knitted

@@
 ---
+
+```{r , include=FALSE}
+
+#installation for the reprtree package.
+options(repos='http://cran.rstudio.org')
+have.packages <- installed.packages()
+cran.packages <- c('devtools','plotrix','randomForest','tree')
+to.install <- setdiff(cran.packages, have.packages[,1])
+if(length(to.install)>0) install.packages(to.install)
+
+library(devtools)
+if(!('reprtree' %in% installed.packages())){
+  install_github('araastat/reprtree')
+}
+for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
+
+```
 
@@
 ```{r SVM_radial}
-SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4)
+train <- read.csv("train.csv")
+test <- read.csv("test.csv")
+train$ChestPainType <- as.factor(train$ChestPainType)
+train$Sex <- as.factor(train$Sex)
+train$FastingBS <- as.factor(train$FastingBS)
+train$RestingECG <- as.factor(train$RestingECG)
+train$ExerciseAngina <- as.factor(train$ExerciseAngina)
+train$ST_Slope <- as.factor(train$ST_Slope)
+test$ChestPainType <- as.factor(test$ChestPainType)
+test$Sex <- as.factor(test$Sex)
+test$FastingBS <- as.factor(test$FastingBS)
+test$RestingECG <- as.factor(test$RestingECG)
+test$ExerciseAngina <- as.factor(test$ExerciseAngina)
+test$ST_Slope <- as.factor(test$ST_Slope)
+
+SVM_radial = svm(HeartDisease~., data=train, kernel="radial", gamma=0.25, cost=0.4,type="C")
 
@@
 
-#table(preds,test$HeartDisease)
+table(preds,test$HeartDisease)
 
@@
 A model like LDA may have one of the highest accuracy, but we could make very little inference about our data from this model.
+
+## References
+
+Banerjee M, Ding Y, Noone AM. Identifying representative trees from ensembles. Stat Med. 2012 Jul 10;31(15):1601-16. doi: 10.1002/sim.4492. Epub 2012 Feb 3. PMID: 22302520.

### 2021-12-16 paper.Rmd
fIXED ECHO

@@
 
-```{r LDA}
+```{r LDA,echo=FALSE}
 LDA_model = lda(HeartDisease~., data=train)
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE,echo=FALSE}
 train_matrix <- model.matrix(HeartDisease ~ ., data = train)
@@
 $d_2(T_1,T_2)=\frac{1}{n}\sum_{i=1}^n(\hat{y}_{1i}-\hat{y}_{2i})^2$, if the trees perfectly similar, the distance $d_2(T_1,T_2)$ is equal 0.
-For for finding the representative tree we used the package "reptree", and using it on this random forests we got the following tree:
+Finding the representative tree was done by using the package "reptree", and using it on this random forests we got the following tree:
 
@@
 
-```{r SVM_linear}
+```{r SVM_linear,echo=FALSE}
 svm_linear_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="linear", data=train,
@@
 
-```{r SVM_poly}
+```{r SVM_poly,echo=FALSE}
 svm_poly_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="polynomial", data=train,
@@
 
-```{r SVM_radial}
+```{r SVM_radial,echo=FALSE}
 train <- read.csv("train.csv")
@@
 
-```{r}
+```{r,echo=FALSE}
 library(knitr)
@@
 
-We found that all of our models, except KNN, had an accuracy within a few points.
+We found that all of our models, except KNN, had an accuracy within a few points as we can see from the table above.
 It was interesting to note from LDA that our prior probabilities of our two classes closely follow the CDCs probability of getting heart disease.
 ST-Slope was shown to be the most important predictor in most of our models and we estimate this is the reason for the similar accuracy between models.
-So choosing the best model depends a lot on whether or not we value more inference or prediction, if we only care about inference, boosting is probably the best method otherwise, something like the logistic regression might be more useful for inference, however, if both are really important a using a random forest with a representative tree might be the best way.
+So choosing the best model depends a lot on whether or not we value more inference or prediction, if we only care about inference, boosting is probably the best method otherwise, something like the logistic regression might be more useful for inference.
+However, if both are really important a using a random forest with a representative tree might be the best way, it allows us to lose some predictive performance from boosting in order to gain the ability to do some interpretation.
 
